---
title: "Figures for adult mosquit abundance under contant and seasonal evironmental conditions"
author: "jbaafi"
date: "2023-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
# Load required packages
pacman::p_load(pacman, deSolve, tidyverse, dplyr, rio)
```

## Plot for Algeria

```{r}
# Clear workspace
rm(list = ls())

# Define a function for the system of equations
pop.model2 <- function(t, y, ...){
  #temp <- 16.93443 - 4.00834*sin(2*pi*t/365) - 10.50303*cos(2*pi*t/365) # Data from Ain Mahbel, Algeria (North Africa)
  temp <- 17 #Algeria average temp
  
  # Monthly average rainfall 
  #R <- 40.31070*(1 + 0.50363*cos(2*pi*(2.34109*t - 275.73821)/365)) # Ain Mahbel, Algeria (Northern Africa)
  R <- 38 #Algeria average rainfall
  
  # Oviposition rate as a function of temperature and rainfall based on Abdelrazec and Gumel, 2017
  #phi_Ao = exp(-0.015*(temp-22)^2)*(1+1.2)*exp(-0.05*(R-10)^2)/(1.2+exp(-0.05*(R-10)^2))
  
  # Oviposition rate (genotrophic cycle) as a function of temperature
  #phi_Ao = exp(-0.015*(temp-22)^2) #  based on Abdelrazec and Gumel, 2017 for Culex mosquitoes
  
  phi_Ao <-  0.443111*exp(-((temp - 34.388928)^2)/185.611107) # based on Beck-Johnson et al, 2013 and Lardeux et al, 2008 on Anopheles mosquito
  
  #phi_Ao <- ifelse(temp <= 15, 0.3973,  -5.4 + 1.8*temp - 0.2124*temp^2 + 0.01015*temp^3 - 0.0001515*temp^4) # Yang et al, 2009 and Wang & Zou, 2018 for Aedes aegypti
  
  # Number of eggs laid per adult female per day (Mordecai et al, 2013, Delatte et al. 2009)
  #b.func <- -0.153*temp^2 + 8.61*temp - 97.7 # species is Aedes albopictus
  #b <- ifelse(temp<=16, 0.1, b.func) 
  
  # Egg hatching/development function based on Abdelrazec & Gumel, 2017
  tauE <- 1.0074 + 6.1694*(1 + (temp/16.6460)^5.8166)^-1  # Average duration spent in egg stage
  mu_E1 <- ifelse(temp <= 15, 0.2415, 
                  ifelse(temp >= 31, 0.2415, 
                         0.0531*exp(((temp - 23)/6.5)^2))) # Egg mortality rate
  PE.T <- exp(-mu_E1) #   daily  survival  probability  of immature  mosquitoes  in  egg stage
  R_IM <- 90 # For algeria data. The maximum rainfall threshold in the community
  P_ME <- 0.9 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PE.R <- R*(R_IM - R)*(4*P_ME / R_IM^2) # The rainfall-depende nt daily probability of survival of mosquito eggs
  PE.TR <- PE.T*PE.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #(PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE)) 
  rho_E <- PE.TR/tauE
  
  # Larvae development rate based on Abdelrazec & Gumel
  tauL <- 7.1361 + 10.1433*(1 + (temp/21.1947)^9.8122)^-1 # Average duration spent in larvae stage
  mu_L1 <- ifelse(temp <= 15, 0.09597, 
                  ifelse(temp >= 31, 0.09597, 
                         0.0211*exp(((temp - 23)/6.5)^2))) # Larvae mortality, Beck-Johnson et al, 2013
  PL.T <- exp(-mu_L1) #   daily  survival  probability  of immature  mosquitoes  in  larvae stage
  P_ML <- 0.9 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PL.R <- R*(R_IM - R)*(4*P_ML / R_IM^2) # The rainfall-dependent daily probability of survival of mosquito eggs
  PL.TR <- PL.T*PL.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #rho_E <- (PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE))  # Development rate of larvae to pupae, Okuneye et al., 
  rho_L <- PL.TR/tauL # Development rate of larvae to pupae Supplemental material by Parham
  
  # Pupa development rate based on Abdelrazec & Gumel
  tauP <- ifelse(temp<=15, 9, 0.9933 + 127.6942*(1 + (temp/5.8999)^2.8134)^-1) # Average duration spent in egg stage
  mu_P1 <- ifelse(temp <= 15, 0.2415, 
                  ifelse(temp >= 31, 0.2415, 
                         0.0531*exp(((temp - 23)/6.5)^2)))
  PP.T <- exp(-mu_P1) #   daily  survival  probability  of immature  mosquitoes  in  egg stage
  P_MP <- 0.90 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PP.R <- R*(R_IM - R)*(4*P_MP / R_IM^2) # The rainfall-dependent daily probability of survival of mosquito eggs
  PP.TR <- PP.T*PP.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #rho_E <- (PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE)) 
  rho_P <- PP.TR/tauP
  
  # Adult mortality rate 
  mu_A <- -log(-0.000828*temp^2 + 0.0367*temp + 0.522) # Taken from Mordecai and Okuneye
  #mu_A <- 0.0005*(temp - 28)^2 + 0.04 # This is from Abdelrezec & Gumel
  
  # The number of states
  E = y[1]
  L = y[2]
  P = y[3]
  Ah = y[4]
  Ar = y[5]
  Ao = y[6]
  
  # The system of equations (This model is based on Lutambi et al, 2013)
  dE <- b*phi_Ao*(1-Ao/K)*Ao - (rho_E + mu_E1 + mu_E2)*E
  dL <- rho_E*E - (rho_L + mu_L1 + mu_L2 + delta_L*L)*L
  dP <- rho_L*L - (rho_P + mu_P1 + mu_P2)*P
  dAh <- rho_P*P + phi_Ao*Ao - (rho_Ah + mu_A + mu_Ah2)*Ah
  dAr <- rho_Ah*Ah - (rho_Ar + mu_A +mu_Ar2)*Ar
  dAo <- rho_Ar*Ar - (phi_Ao + mu_A + mu_Ao2)*Ao
  #return(list(c(dE, dL, dP, dAh, dAr, dAo, temp, R)))
  return(list(c(dE, dL, dP, dAh, dAr, dAo)))
}

# Model parameters (All parameter values are based on Lutambi et al, 2013) and adapted to Anopheles species
b  = 200      # number of eggs laid per oviposition (range  100 - 400 eggs at a time) https://www.canada.ca/en/health-canada/services/pest-control-tips/mosquitoes.html
mu_E2 = 0.005    # extra natural mortality of eggs that is not dependent on environmental conditions
mu_L2 = 0.005 # Natural mortality of larvae
delta_L = 0.01 # For every 100 larvae, 1 larvae will probably die due to competition for resources (Assumed)
mu_P2 = 0.005 # Natural pupal mortality
rho_Ah =  2 # 0.46 # 2 (According to Cailly et al, 2012, this parameter takes the value of 2 day^-1) 
mu_Ah2 = 0.01 
rho_Ar =  1/2 # Takes 2 days for egg to develop and seek oviposition site
mu_Ar2 = 0.005 
mu_Ao2 = 0.01
K = 10^13  # From Okuneye et al, 

param <- c(b, 
           mu_E2, 
           mu_L2, 
           delta_L, 
           mu_P2, 
           rho_Ah, 
           mu_Ah2, 
           rho_Ar, 
           mu_Ar2, 
           mu_Ao2, 
           K)

# Initial conditions (I am not sure what values to use as initial conditions but I guess it doesn't matter much)
# y.initial <- c(100.0, 100.0, 100.0, 30.0, 30.0, 40.0, 6.4314, 41.0054)
y.initial <- c(100.0, 100.0, 100.0, 30.0, 30.0, 40.0)

# Time steps
start <- 0
end <- 365*5
times <- seq(start, end, by = 1)


# Numerical integration using the ode() function from deSolve package
out <-  ode(y = y.initial, func = pop.model2, times = times, parms = param, method = "lsoda") %>% 
  data.frame()

# Rename variable
# out2 <- out2 %>%
#   rename(eggs = X1, larvae = X2, pupae = X3, host.seeking.adults = X4,
#          resting.adults = X5, ovipositn.adults = X6, temp=X7, rain=X8)

out <- out %>%
  rename(eggs = X1, larvae = X2, pupae = X3, host.seeking.adults = X4,
         resting.adults = X5, ovipositn.adults = X6)

#Add a column which contains the total number of adults (A = Ah + Ar + Ao) to the data.frame
out <- out %>%
  mutate(fixed.total.adults =  out$host.seeking.adults + out$resting.adults + out$ovipositn.adults)
#

# out <- out %>%
#   mutate(vary.total.adults =  out$host.seeking.adults + out$resting.adults + out$ovipositn.adults)
# 

# Temp and rain need to be differentiated as it gets integrated through the function
# out <- out %>%
#   mutate(temperature =  c(out$temp[1],diff(out$temp)*(1/1)),
#          rainfall = c(out$rain[1],diff(out$rain)*(1/1)))

# This is when time step is 0.1. I changed it to 1 so we don't need this part.
#out2 <- out2 %>% 
#   mutate(temperature =  c(out2$temp[1],diff(out2$temp)*(1/0.1)), 
#         rainfall = c(out2$rain[1],diff(out2$rain)*(1/0.1)))

# I wanna drop temp and rain column
# df <- out2 %>%
#   subset(select = -c(temp, rain) )

#data <- write_csv(df, "data.csv")

df <- out
# # Define a function to produce the plots
visualize <- function(p){
  # Set plot environment
  par(mfrow = c(3, 2), mar=c(2,4,2,2))    # 2 rows, 2 columns
  
  # Create plots
  plot(df$time, df$eggs, type = "l", xlab = "Time (days)", ylab = "Egg abundance",
       pch = 19, col = "blue")  # plot no. 1
  
  plot(df$time, df$larvae, type = "l", xlab = "Time (days)", ylab = "Larval abundance",
       pch = 19, col = "green")     # plot no. 2
  
  plot(df$time, df$pupae, type = "l", xlab = "Time (days)", ylab = "Pupal abundance",
       pch = 19, col = "yellow")     # plot no. 3
  
  plot(df$time, df$host.seeking.adults, type = "l", xlab = "Time (days)", ylab = "Adults seeking hosts",
       pch = 19, col = "red")    # plot no. 4
  
  plot(df$time, df$resting.adults, type = "l", xlab = "Time (days)", ylab = "Adults resting",
       pch = 19, col = "violet")    # plot no. 5
  
  plot(df$time, df$ovipositn.adults, type = "l", xlab = "Time (days)", ylab = "Adults ovipositing",
       pch = 19, col = "orange")    # plot no. 6
}

visualize(p)

fixed.condt_Alg <- df 
#variable.condt_Alg <- df

#variable.condt_Alg <- df 
#write_csv(fixed.condt_Alg, file = "/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_Alg.csv")
#write_csv(variable.condt_Alg, file = "/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_Alg.csv")

##- --------------------------------------------------------------------------------------------------
#. I use saved data from the above simulation analysis to produce figures for Algeria in the next chunk.
##- --------------------------------------------------------------------------------------------------
```


```{r}
# This will always run because I have saved the data and does not depend on the previous simulation

# Clear space
rm(list = ls())

# Load data (on macbook)
df_fix <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_Alg.csv")
df_vary <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_Alg.csv")

# Load data (on mac-mini)
#df_fix <- read.csv("/Users/josephbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_Alg.csv")
#df_vary <- read.csv("/Users/josephbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_Alg.csv")

# Select needed columns
df.fix <- df_fix %>% 
  select(time, fixed.total.adults)

# Select needed columns
df.vary <- df_vary %>% 
  select(time, vary.total.adults)

# Put together into one dataframe
df.alg.combined <- data.frame(time = df.fix$time, constant = df.fix$fixed.total.adults,
                              seasonal = df.vary$vary.total.adults)

# Select the year 5 data 
df.alg.combined <- df.alg.combined[1461:nrow(df.alg.combined), ]

# Convert data to the long format
df.alg.combined <- df.alg.combined %>% 
  pivot_longer(cols=c("constant", "seasonal"),
               names_to='Environment',
               values_to='value')

# Plot data 
population.alg <- ggplot(data = df.alg.combined, aes(x=time, y=value, col = Environment))+
  geom_line(linewidth = 1)+
  xlab("Time (days)") +
  ylab("Population") +
  theme_classic()

population.alg

# rename columns
df.alg.combined2 <- df.alg.combined %>% 
  rename(
    x = time,
    y = value
  )

# Change time to months
breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df.alg.combined2$y) * 1.1  # Adjust the buffer as needed (e.g., 1.1 for a 10% buffer)


ggplot(df.alg.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "Algeria",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.8),  # Adjust the values to move the legend
    legend.box = "horizontal",  # Use "vertical" for a vertical box
    legend.box.background = element_rect(color = "black", size = 1)  # Rectangle around the legend
  ) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit))


# Define the colors you want to use
line_colors <- c("red", "purple")

# The plot
plot.alg <- ggplot(df.alg.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "Algeria",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.8)
  ) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit)) +
  scale_color_manual(values = line_colors)

plot.alg

# Save plot
#ggsave(plot.alg, filename = "constant_and_seasonal_env_algeria.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```


I include a chunk here to do a few calculations that is to be added to the manuscripts
```{r}
## Calculate total number of mosquito for constant and seasonal environments
data_constant <- df.alg.combined2 %>% 
  filter(Environment == "constant")

constant_pop_abundance <- sum(data_constant$y)
constant_pop_abundance

###
data_seasonal <- df.alg.combined2 %>% 
  filter(Environment == "seasonal")

seasonal_pop_abudance <- sum(data_seasonal$y)
seasonal_pop_abudance

####
seasonal_pop_abudance/constant_pop_abundance
```


## Plot for Ghana

```{r}
# Define a function for the system of equations
pop.model2 <- function(t, y, ...){
  # Temperature
  #temp <- 27.36355 + 1.79244*sin(2*pi*t/365) + 1.43960*cos(2*pi*t/365) # Data from Kumasi Ghana (West Africa)
  temp <- 27 #Ghana average temp

  # Monthly average rainfall 
  #R <- 195.648 - 33.234*sin(2*pi*t/365) - 112.033*cos(2*pi*t/365) # Kumasi, Ghana
  R <- 195 #Ghana average rainfall
 
  # Oviposition rate as a function of temperature and rainfall based on Abdelrazec and Gumel, 2017
  #phi_Ao = exp(-0.015*(temp-22)^2)*(1+1.2)*exp(-0.05*(R-10)^2)/(1.2+exp(-0.05*(R-10)^2))
  
  # Oviposition rate (genotrophic cycle) as a function of temperature
  #phi_Ao = exp(-0.015*(temp-22)^2) #  based on Abdelrazec and Gumel, 2017 for Culex mosquitoes
  
  phi_Ao <-  0.443111*exp(-((temp - 34.388928)^2)/185.611107) # based on Beck-Johnson et al, 2013 and Lardeux et al, 2008 on Anopheles mosquito
  
  #phi_Ao <- ifelse(temp <= 15, 0.3973,  -5.4 + 1.8*temp - 0.2124*temp^2 + 0.01015*temp^3 - 0.0001515*temp^4) # Yang et al, 2009 and Wang & Zou, 2018 for Aedes aegypti
  
  # Egg hatching/development function based on Abdelrazec & Gumel, 2017
  tauE <- 1.0074 + 6.1694*(1 + (temp/16.6460)^5.8166)^-1  # Average duration spent in egg stage
  mu_E1 <- ifelse(temp <= 15, 0.2415, 
                  ifelse(temp >= 31, 0.2415, 
                         0.0531*exp(((temp - 23)/6.5)^2))) # Egg mortality rate
  PE.T <- exp(-mu_E1) #   daily  survival  probability  of immature  mosquitoes  in  egg stage
  R_IM <- 350 # For Ghana data
  P_ME <- 0.9 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PE.R <- R*(R_IM - R)*(4*P_ME / R_IM^2) # The rainfall-depende nt daily probability of survival of mosquito eggs
  PE.TR <- PE.T*PE.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #(PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE)) 
  rho_E <- PE.TR/tauE
  
  
  # Larvae development rate based on Abdelrazec & Gumel
  tauL <- 7.1361 + 10.1433*(1 + (temp/21.1947)^9.8122)^-1 # Average duration spent in larvae stage
  mu_L1 <- ifelse(temp <= 15, 0.09597, 
                  ifelse(temp >= 31, 0.09597, 
                         0.0211*exp(((temp - 23)/6.5)^2))) # Larvae mortality, Beck-Johnson et al, 2013
  PL.T <- exp(-mu_L1) #   daily  survival  probability  of immature  mosquitoes  in  larvae stage
  P_ML <- 0.9 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PL.R <- R*(R_IM - R)*(4*P_ML / R_IM^2) # The rainfall-dependent daily probability of survival of mosquito eggs
  PL.TR <- PL.T*PL.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #rho_E <- (PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE))  # Development rate of larvae to pupae, Okuneye et al., 
  rho_L <- PL.TR/tauL # Development rate of larvae to pupae Supplemental material by Parham
  
  # Pupa development rate based on Abdelrazec & Gumel
  tauP <- ifelse(temp<=15, 9, 0.9933 + 127.6942*(1 + (temp/5.8999)^2.8134)^-1) # Average duration spent in egg stage
  mu_P1 <- ifelse(temp <= 15, 0.2415, 
                  ifelse(temp >= 31, 0.2415, 
                         0.0531*exp(((temp - 23)/6.5)^2)))
  PP.T <- exp(-mu_P1) #   daily  survival  probability  of immature  mosquitoes  in  egg stage
  P_MP <- 0.90 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PP.R <- R*(R_IM - R)*(4*P_MP / R_IM^2) # The rainfall-dependent daily probability of survival of mosquito eggs
  PP.TR <- PP.T*PP.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #rho_E <- (PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE)) 
  rho_P <- PP.TR/tauP
  
  # Adult mortality rate 
  mu_A <- -log(-0.000828*temp^2 + 0.0367*temp + 0.522) # Taken from Mordecai and Okuneye
  #mu_A <- 0.0005*(temp - 28)^2 + 0.04 # This is from Abdelrezec & Gumel
  
  # The number of states
  E = y[1]
  L = y[2]
  P = y[3]
  Ah = y[4]
  Ar = y[5]
  Ao = y[6]
  
  # The system of equations (This model is based on Lutambi et al, 2013)
  dE <- b*phi_Ao*(1-Ao/K)*Ao - (rho_E + mu_E1 + mu_E2)*E
  dL <- rho_E*E - (rho_L + mu_L1 + mu_L2 + delta_L*L)*L
  dP <- rho_L*L - (rho_P + mu_P1 + mu_P2)*P
  dAh <- rho_P*P + phi_Ao*Ao - (rho_Ah + mu_A + mu_Ah2)*Ah
  dAr <- rho_Ah*Ah - (rho_Ar + mu_A +mu_Ar2)*Ar
  dAo <- rho_Ar*Ar - (phi_Ao + mu_A + mu_Ao2)*Ao
  #return(list(c(dE, dL, dP, dAh, dAr, dAo, temp, R)))
  return(list(c(dE, dL, dP, dAh, dAr, dAo)))
}

# Model parameters (All parameter values are based on Lutambi et al, 2013) and adapted to Anopheles species
b  = 200      # number of eggs laid per oviposition (range  100 - 400 eggs at a time) https://www.canada.ca/en/health-canada/services/pest-control-tips/mosquitoes.html
mu_E2 = 0.005    # extra natural mortality of eggs that is not dependent on environmental conditions
mu_L2 = 0.005 # Natural mortality of larvae
delta_L = 0.01 # For every 100 larvae, 1 larvae will probably die due to competition for resources (Assumed)
mu_P2 = 0.005 # Natural pupal mortality
rho_Ah =  2 # 0.46 # 2 (According to Cailly et al, 2012, this parameter takes the value of 2 day^-1) 
mu_Ah2 = 0.01 
rho_Ar =  1/2 # Takes 2 days for egg to develop and seek oviposition site
mu_Ar2 = 0.005 
mu_Ao2 = 0.01
K = 10^13  # From Okuneye et al, 

param <- c(b, 
           mu_E2, 
           mu_L2, 
           delta_L, 
           mu_P2, 
           rho_Ah, 
           mu_Ah2, 
           rho_Ar, 
           mu_Ar2, 
           mu_Ao2, 
           K)

# Initial conditions (I am not sure what values to use as initial conditions but I guess it doesn't matter much)
#y.initial <- c(100.0, 100.0, 100.0, 30.0, 30.0, 40.0, 6.4314, 41.0054)
y.initial <- c(100.0, 100.0, 100.0, 30.0, 30.0, 40.0)

# Time steps (Period with which to run the model)
start <- 0
end <- 365*5
times <- seq(start, end, by = 1)

# Numerical integration using the ode() function from deSolve package
out2 <-  ode(y = y.initial, func = pop.model2, times = times, parms = param, method = "lsoda") %>% 
  data.frame()

# Rename variable
# out2 <- out2 %>%
#   rename(eggs = X1, larvae = X2, pupae = X3, host.seeking.adults = X4,
#          resting.adults = X5, ovipositn.adults = X6, temp=X7, rain=X8)

out2 <- out2 %>%
  rename(eggs = X1, larvae = X2, pupae = X3, host.seeking.adults = X4,
         resting.adults = X5, ovipositn.adults = X6)

#Add a column which contains the total number of adults (A = Ah + Ar + Ao) to the data.frame
out2 <- out2 %>%
  mutate(fixed.total.adults =  out2$host.seeking.adults + out2$resting.adults + out2$ovipositn.adults)
#
# 
# out2 <- out2 %>%
#   mutate(vary.total.adults =  out2$host.seeking.adults + out2$resting.adults + out2$ovipositn.adults)
# 

# Temp and rain need to be differentiated as it gets integrated through the function
# out2 <- out2 %>% 
#   mutate(temperature =  c(out2$temp[1],diff(out2$temp)*(1/1)), 
#          rainfall = c(out2$rain[1],diff(out2$rain)*(1/1)))
# 
# This is when time step is 0.1. I changed it to 1 so we don't need this part.
#out2 <- out2 %>% 
#   mutate(temperature =  c(out2$temp[1],diff(out2$temp)*(1/0.1)), 
#         rainfall = c(out2$rain[1],diff(out2$rain)*(1/0.1)))

# I wanna drop temp and rain column
# df <- out2 %>%
#   subset(select = -c(temp, rain) )

#data <- write_csv(df, "data.csv")

df <- out2
# # Define a function to produce the plots
visualize <- function(p){
  # Set plot environment
  par(mfrow = c(3, 2), mar=c(2,4,2,2))    # 2 rows, 2 columns
  
  # Create plots
  plot(df$time, df$eggs, type = "l", xlab = "Time (days)", ylab = "Egg abundance",
       pch = 19, col = "blue")  # plot no. 1
  
  plot(df$time, df$larvae, type = "l", xlab = "Time (days)", ylab = "Larval abundance",
       pch = 19, col = "green")     # plot no. 2
  
  plot(df$time, df$pupae, type = "l", xlab = "Time (days)", ylab = "Pupal abundance",
       pch = 19, col = "yellow")     # plot no. 3
  
  plot(df$time, df$host.seeking.adults, type = "l", xlab = "Time (days)", ylab = "Adults seeking hosts",
       pch = 19, col = "red")    # plot no. 4
  
  plot(df$time, df$resting.adults, type = "l", xlab = "Time (days)", ylab = "Adults resting",
       pch = 19, col = "violet")    # plot no. 5
  
  plot(df$time, df$ovipositn.adults, type = "l", xlab = "Time (days)", ylab = "Adults ovipositing",
       pch = 19, col = "orange")    # plot no. 6
}

visualize(p)

fixed.condt_gh <- df 
#variable.condt_gh <- df 

#write_csv(fixed.condt_gh, file = "/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_gh.csv")
#write_csv(variable.condt_gh, file = "/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")

```


```{r}
##- --------------------------------------------------------------------------------------------------
#. I use saved data from the above simulation analysis to produce figures for Ghana
##- --------------------------------------------------------------------------------------------------

# Load data (on macbook)
df_fix.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_gh.csv")
df_vary.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")

# Load data (on macmini)
# df_fix.gh <- read.csv("/Users/josephbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_gh.csv")
# df_vary.gh <- read.csv("/Users/josephbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")

df.fix.gh <- df_fix.gh %>% 
  select(time, fixed.total.adults)

df.vary.gh <- df_vary.gh %>% 
  select(time, vary.total.adults)


df.gh.combined <- data.frame(time = df.fix.gh$time, constant = df.fix.gh$fixed.total.adults,
                              seasonal = df.vary.gh$vary.total.adults)

df.gh.combined <- df.gh.combined[1460:nrow(df.gh.combined), ]

df.gh.combined <- df.gh.combined %>% 
  pivot_longer(cols=c("constant", "seasonal"),
               names_to='Environment',
               values_to='value')

# Plot data 
population.gh <- ggplot(data = df.gh.combined, aes(x=time, y=value, col = Environment))+
  geom_line(linewidth = 1)+
  xlab("Time (days)") +
  ylab("Population") +
  theme_classic()

population.gh

# rename columns
df.gh.combined2 <- df.gh.combined %>% 
  rename(
    x = time,
    y = value
  )

# Change time to months
breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df.gh.combined2$y) * 1.1  # Adjust the buffer as needed (e.g., 1.1 for a 10% buffer)

ggplot(df.gh.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "Ghana",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.7),  # Adjust the values to move the legend
    legend.box = "horizontal",  # Use "vertical" for a vertical box
    legend.box.background = element_rect(color = "black", size = 1)  # Rectangle around the legend
  ) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit))


# Define the colors you want to use
line_colors <- c("red", "darkred")

# The plot
plot.gh <- ggplot(df.gh.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "Ghana",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.7)) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit)) +
scale_color_manual(values = line_colors)

plot.gh
  
# Save plot
#ggsave(plot.gh, filename = "constant_and_seasonal_env_ghana.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```

some calculations to be included in the manuscript
```{r}
data_seasonal_gh <- df.gh.combined2 %>% 
  filter(Environment == "seasonal")

seasonal_pop_abudance_gh <- sum(data_seasonal_gh$y)
seasonal_pop_abudance_gh
```


## Plot for Kenya

```{r}
# Define a function for the system of equations
pop.model2 <- function(t, y, ...){
  # A function for daily temperature 
  temp <- 19.86656 + 0.32835*sin(2*pi*t/365) + 0.83746*cos(2*pi*t/365) # Data from Nairobi, Kenya (East Africa) (Not too good fit)
  #temp <- 20 #Kenya average temp
  
  # Monthly average rainfall 
  R <- 114.94298*(1 + 0.52783*cos(2*pi*(1.69070*t - 199.83939)/365)) # Nairobi, Kenya (East Africa)
  #R <- 118 #Kenya average rainfall
  
  # Oviposition rate as a function of temperature and rainfall based on Abdelrazec and Gumel, 2017
  #phi_Ao = exp(-0.015*(temp-22)^2)*(1+1.2)*exp(-0.05*(R-10)^2)/(1.2+exp(-0.05*(R-10)^2))
  
  # Oviposition rate (genotrophic cycle) as a function of temperature
  #phi_Ao = exp(-0.015*(temp-22)^2) #  based on Abdelrazec and Gumel, 2017 for Culex mosquitoes
  
  phi_Ao <-  0.443111*exp(-((temp - 34.388928)^2)/185.611107) # based on Beck-Johnson et al, 2013 and Lardeux et al, 2008 on Anopheles mosquito
  
  #phi_Ao <- ifelse(temp <= 15, 0.3973,  -5.4 + 1.8*temp - 0.2124*temp^2 + 0.01015*temp^3 - 0.0001515*temp^4) # Yang et al, 2009 and Wang & Zou, 2018 for Aedes aegypti
  
  # Number of eggs laid per adult female per day (Mordecai et al, 2013, Delatte et al. 2009)
  #b.func <- -0.153*temp^2 + 8.61*temp - 97.7 # species is Aedes albopictus
  #b <- ifelse(temp<=16, 0.1, b.func) 
  
  # Egg hatching/development function based on Abdelrazec & Gumel, 2017
  tauE <- 1.0074 + 6.1694*(1 + (temp/16.6460)^5.8166)^-1  # Average duration spent in egg stage
  mu_E1 <- ifelse(temp <= 15, 0.2415, 
                  ifelse(temp >= 31, 0.2415, 
                         0.0531*exp(((temp - 23)/6.5)^2))) # Egg mortality rate
  PE.T <- exp(-mu_E1) #   daily  survival  probability  of immature  mosquitoes  in  egg stage
  R_IM <- 220 # For Kenya data
  P_ME <- 0.9 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PE.R <- R*(R_IM - R)*(4*P_ME / R_IM^2) # The rainfall-depende nt daily probability of survival of mosquito eggs
  PE.TR <- PE.T*PE.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #(PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE)) 
  rho_E <- PE.TR/tauE
  
  
  # Larvae development rate based on Abdelrazec & Gumel
  tauL <- 7.1361 + 10.1433*(1 + (temp/21.1947)^9.8122)^-1 # Average duration spent in larvae stage
  mu_L1 <- ifelse(temp <= 15, 0.09597, 
                  ifelse(temp >= 31, 0.09597, 
                         0.0211*exp(((temp - 23)/6.5)^2))) # Larvae mortality, Beck-Johnson et al, 2013
  PL.T <- exp(-mu_L1) #   daily  survival  probability  of immature  mosquitoes  in  larvae stage
  P_ML <- 0.9 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PL.R <- R*(R_IM - R)*(4*P_ML / R_IM^2) # The rainfall-dependent daily probability of survival of mosquito eggs
  PL.TR <- PL.T*PL.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #rho_E <- (PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE))  # Development rate of larvae to pupae, Okuneye et al., 
  rho_L <- PL.TR/tauL # Development rate of larvae to pupae Supplemental material by Parham
  
  # Pupa development rate based on Abdelrazec & Gumel
  tauP <- ifelse(temp<=15, 9, 0.9933 + 127.6942*(1 + (temp/5.8999)^2.8134)^-1) # Average duration spent in egg stage
  mu_P1 <- ifelse(temp <= 15, 0.2415, 
                  ifelse(temp >= 31, 0.2415, 
                         0.0531*exp(((temp - 23)/6.5)^2)))
  PP.T <- exp(-mu_P1) #   daily  survival  probability  of immature  mosquitoes  in  egg stage
  P_MP <- 0.90 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PP.R <- R*(R_IM - R)*(4*P_MP / R_IM^2) # The rainfall-dependent daily probability of survival of mosquito eggs
  PP.TR <- PP.T*PP.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #rho_E <- (PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE)) 
  rho_P <- PP.TR/tauP
  
  # Adult mortality rate 
  mu_A <- -log(-0.000828*temp^2 + 0.0367*temp + 0.522) # Taken from Mordecai and Okuneye
  #mu_A <- 0.0005*(temp - 28)^2 + 0.04 # This is from Abdelrezec & Gumel
  
  # The number of states
  E = y[1]
  L = y[2]
  P = y[3]
  Ah = y[4]
  Ar = y[5]
  Ao = y[6]
  
  # The system of equations (This model is based on Lutambi et al, 2013)
  dE <- b*phi_Ao*(1-Ao/K)*Ao - (rho_E + mu_E1 + mu_E2)*E
  dL <- rho_E*E - (rho_L + mu_L1 + mu_L2 + delta_L*L)*L
  dP <- rho_L*L - (rho_P + mu_P1 + mu_P2)*P
  dAh <- rho_P*P + phi_Ao*Ao - (rho_Ah + mu_A + mu_Ah2)*Ah
  dAr <- rho_Ah*Ah - (rho_Ar + mu_A +mu_Ar2)*Ar
  dAo <- rho_Ar*Ar - (phi_Ao + mu_A + mu_Ao2)*Ao
  #return(list(c(dE, dL, dP, dAh, dAr, dAo, temp, R)))
  return(list(c(dE, dL, dP, dAh, dAr, dAo)))
}

# Model parameters (All parameter values are based on Lutambi et al, 2013) and adapted to Anopheles species
b  = 200      # number of eggs laid per oviposition (range  100 - 400 eggs at a time) https://www.canada.ca/en/health-canada/services/pest-control-tips/mosquitoes.html
mu_E2 = 0.005    # extra natural mortality of eggs that is not dependent on environmental conditions
mu_L2 = 0.005 # Natural mortality of larvae
delta_L = 0.01 # For every 100 larvae, 1 larvae will probably die due to competition for resources (Assumed)
mu_P2 = 0.005 # Natural pupal mortality
rho_Ah =  2 # 0.46 # 2 (According to Cailly et al, 2012, this parameter takes the value of 2 day^-1) 
mu_Ah2 = 0.01 
rho_Ar =  1/2 # Takes 2 days for egg to develop and seek oviposition site
mu_Ar2 = 0.005 
mu_Ao2 = 0.01
K = 10^13  # From Okuneye et al, 

param <- c(b, 
           mu_E2, 
           mu_L2, 
           delta_L, 
           mu_P2, 
           rho_Ah, 
           mu_Ah2, 
           rho_Ar, 
           mu_Ar2, 
           mu_Ao2, 
           K)

# Initial conditions (I am not sure what values to use as initial conditions but I guess it doesn't matter much)
#y.initial <- c(100.0, 100.0, 100.0, 30.0, 30.0, 40.0, 6.4314, 41.0054)
y.initial <- c(100.0, 100.0, 100.0, 30.0, 30.0, 40.0)

# Time steps (Period with which to run the model)

start <- 0
end <- 365*5
times <- seq(start, end, by = 1)


# Numerical integration using the ode() function from deSolve package
out2 <-  ode(y = y.initial, func = pop.model2, times = times, parms = param, method = "lsoda") %>% 
  data.frame()

# Rename variable
# out2 <- out2 %>%
#   rename(eggs = X1, larvae = X2, pupae = X3, host.seeking.adults = X4,
#          resting.adults = X5, ovipositn.adults = X6, temp=X7, rain=X8)

out2 <- out2 %>%
  rename(eggs = X1, larvae = X2, pupae = X3, host.seeking.adults = X4,
         resting.adults = X5, ovipositn.adults = X6)

#Add a column which contains the total number of adults (A = Ah + Ar + Ao) to the data.frame
# out2 <- out2 %>% 
#   mutate(fixed.total.adults =  out2$host.seeking.adults + out2$resting.adults + out2$ovipositn.adults)
# # 

out2 <- out2 %>%
  mutate(vary.total.adults =  out2$host.seeking.adults + out2$resting.adults + out2$ovipositn.adults)

# Temp and rain need to be differentiated as it gets integrated through the function
# out2 <- out2 %>% 
#   mutate(temperature =  c(out2$temp[1],diff(out2$temp)*(1/1)), 
#          rainfall = c(out2$rain[1],diff(out2$rain)*(1/1)))

# This is when time step is 0.1. I changed it to 1 so we don't need this part.
#out2 <- out2 %>% 
#   mutate(temperature =  c(out2$temp[1],diff(out2$temp)*(1/0.1)), 
#         rainfall = c(out2$rain[1],diff(out2$rain)*(1/0.1)))

# # I wanna drop temp and rain column
# df <- out2 %>%
#   subset(select = -c(temp, rain) )

#data <- write_csv(df, "data.csv")

df <- out2
# # Define a function to produce the plots
visualize <- function(p){
  # Set plot environment
  par(mfrow = c(3, 2), mar=c(2,4,2,2))    # 2 rows, 2 columns
  
  # Create plots
  plot(df$time, df$eggs, type = "l", xlab = "Time (days)", ylab = "Egg abundance",
       pch = 19, col = "blue")  # plot no. 1
  
  plot(df$time, df$larvae, type = "l", xlab = "Time (days)", ylab = "Larval abundance",
       pch = 19, col = "green")     # plot no. 2
  
  plot(df$time, df$pupae, type = "l", xlab = "Time (days)", ylab = "Pupal abundance",
       pch = 19, col = "yellow")     # plot no. 3
  
  plot(df$time, df$host.seeking.adults, type = "l", xlab = "Time (days)", ylab = "Adults seeking hosts",
       pch = 19, col = "red")    # plot no. 4
  
  plot(df$time, df$resting.adults, type = "l", xlab = "Time (days)", ylab = "Adults resting",
       pch = 19, col = "violet")    # plot no. 5
  
  plot(df$time, df$ovipositn.adults, type = "l", xlab = "Time (days)", ylab = "Adults ovipositing",
       pch = 19, col = "orange")    # plot no. 6
}

visualize(p)

#fixed.condt_kenya <- df 
variable.condt_kenya <- df 

#write_csv(fixed.condt_kenya, file = "/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_kenya.csv")
#write_csv(variable.condt_kenya, file = "/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")
```


```{r}

##- --------------------------------------------------------------------------------------------------
#.  Figures for Kenya
##- --------------------------------------------------------------------------------------------------

# Load data
df_fix.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_kenya.csv")
df_vary.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")

df.fix.kenya <- df_fix.kenya %>% 
  select(time, fixed.total.adults)

df.vary.kenya <- df_vary.kenya %>% 
  select(time, vary.total.adults)


df.kenya.combined <- data.frame(time = df.fix.kenya$time, constant = df.fix.kenya$fixed.total.adults,
                             seasonal = df.vary.kenya$vary.total.adults)

df.kenya.combined <- df.kenya.combined[1460:nrow(df.kenya.combined), ]

df.kenya.combined <- df.kenya.combined %>% 
  pivot_longer(cols=c("constant", "seasonal"),
               names_to='Environment',
               values_to='value')

# Plot data 
population.kenya <- ggplot(data = df.kenya.combined, aes(x=time, y=value, col = Environment))+
  geom_line(linewidth = 1)+
  xlab("Time (days)") +
  ylab("Population") +
  theme_classic()

population.kenya

# rename columns
df.kenya.combined2 <- df.kenya.combined %>% 
  rename(
    x = time,
    y = value
  )

# Change time to months
breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df.kenya.combined2$y) * 1.1  # Adjust the buffer as needed (e.g., 1.1 for a 10% buffer)

ggplot(df.kenya.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "Kenya",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.4),  # Adjust the values to move the legend
    legend.box = "horizontal",  # Use "vertical" for a vertical box
    legend.box.background = element_rect(color = "black", size = 1)  # Rectangle around the legend
  ) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit))


# Define the colors you want to use
line_colors <- c("red", "orange")

plot.kenya <- ggplot(df.kenya.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "Kenya",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.4)) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit)) +
scale_color_manual(values = line_colors)

plot.kenya
# Save 
#ggsave(plot.kenya, filename = "constant_and_seasonal_env_kenya.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```



## Plot for South Africa

```{r}
# Define a function for the system of equations
pop.model2 <- function(t, y, ...){
  # A function for daily temperature
  
  temp <- 18.01768 + 2.02738*sin(2*pi*t/365) + 3.53049*cos(2*pi*t/365) # Data from Cape Town, SA (Good fit)
  #temp <- 18 #SA average temp
  
  
  # Monthly average rainfall 
  R <- 76.891 - 9.785*sin(2*pi*t/365) - 71.288*cos(2*pi*t/365) # Cape Town, South Africa (Southern Africa)
  #R <- 77 #SA average rainfall
  
  # Oviposition rate as a function of temperature and rainfall based on Abdelrazec and Gumel, 2017
  #phi_Ao = exp(-0.015*(temp-22)^2)*(1+1.2)*exp(-0.05*(R-10)^2)/(1.2+exp(-0.05*(R-10)^2))
  
  # Oviposition rate (genotrophic cycle) as a function of temperature
  #phi_Ao = exp(-0.015*(temp-22)^2) #  based on Abdelrazec and Gumel, 2017 for Culex mosquitoes
  
  phi_Ao <-  0.443111*exp(-((temp - 34.388928)^2)/185.611107) # based on Beck-Johnson et al, 2013 and Lardeux et al, 2008 on Anopheles mosquito
  
  #phi_Ao <- ifelse(temp <= 15, 0.3973,  -5.4 + 1.8*temp - 0.2124*temp^2 + 0.01015*temp^3 - 0.0001515*temp^4) # Yang et al, 2009 and Wang & Zou, 2018 for Aedes aegypti
  
  # Number of eggs laid per adult female per day (Mordecai et al, 2013, Delatte et al. 2009)
  #b.func <- -0.153*temp^2 + 8.61*temp - 97.7 # species is Aedes albopictus
  #b <- ifelse(temp<=16, 0.1, b.func) 
  
  # Egg hatching/development function based on Abdelrazec & Gumel, 2017
  tauE <- 1.0074 + 6.1694*(1 + (temp/16.6460)^5.8166)^-1  # Average duration spent in egg stage
  mu_E1 <- ifelse(temp <= 15, 0.2415, 
                  ifelse(temp >= 31, 0.2415, 
                         0.0531*exp(((temp - 23)/6.5)^2))) # Egg mortality rate
  PE.T <- exp(-mu_E1) #   daily  survival  probability  of immature  mosquitoes  in  egg stage
  R_IM <- 200 # For SA data
  P_ME <- 0.9 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PE.R <- R*(R_IM - R)*(4*P_ME / R_IM^2) # The rainfall-depende nt daily probability of survival of mosquito eggs
  PE.TR <- PE.T*PE.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #(PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE)) 
  rho_E <- PE.TR/tauE
  
  # Larvae development rate based on Abdelrazec & Gumel
  tauL <- 7.1361 + 10.1433*(1 + (temp/21.1947)^9.8122)^-1 # Average duration spent in larvae stage
  mu_L1 <- ifelse(temp <= 15, 0.09597, 
                  ifelse(temp >= 31, 0.09597, 
                         0.0211*exp(((temp - 23)/6.5)^2))) # Larvae mortality, Beck-Johnson et al, 2013
  PL.T <- exp(-mu_L1) #   daily  survival  probability  of immature  mosquitoes  in  larvae stage
  P_ML <- 0.9 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PL.R <- R*(R_IM - R)*(4*P_ML / R_IM^2) # The rainfall-dependent daily probability of survival of mosquito eggs
  PL.TR <- PL.T*PL.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #rho_E <- (PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE))  # Development rate of larvae to pupae, Okuneye et al., 
  rho_L <- PL.TR/tauL # Development rate of larvae to pupae Supplemental material by Parham
  
  # Pupa development rate based on Abdelrazec & Gumel
  tauP <- ifelse(temp<=15, 9, 0.9933 + 127.6942*(1 + (temp/5.8999)^2.8134)^-1) # Average duration spent in egg stage
  mu_P1 <- ifelse(temp <= 15, 0.2415, 
                  ifelse(temp >= 31, 0.2415, 
                         0.0531*exp(((temp - 23)/6.5)^2)))
  PP.T <- exp(-mu_P1) #   daily  survival  probability  of immature  mosquitoes  in  egg stage
  P_MP <- 0.90 # Peak  daily  survival  probability  of  immature  mosquito  egg stage
  PP.R <- R*(R_IM - R)*(4*P_MP / R_IM^2) # The rainfall-dependent daily probability of survival of mosquito eggs
  PP.TR <- PP.T*PP.R #  The  daily  survival  probability  of immature mosquitoes  in egg stage
  #rho_E <- (PE.TR^tauE)*(1 - PE.TR)/(1 - (PE.TR^tauE)) 
  rho_P <- PP.TR/tauP
  
  # Adult mortality rate 
  mu_A <- -log(-0.000828*temp^2 + 0.0367*temp + 0.522) # Taken from Mordecai and Okuneye
  #mu_A <- 0.0005*(temp - 28)^2 + 0.04 # This is from Abdelrezec & Gumel
  
  # The number of states
  E = y[1]
  L = y[2]
  P = y[3]
  Ah = y[4]
  Ar = y[5]
  Ao = y[6]
  
  # The system of equations (This model is based on Lutambi et al, 2013)
  dE <- b*phi_Ao*(1-Ao/K)*Ao - (rho_E + mu_E1 + mu_E2)*E
  dL <- rho_E*E - (rho_L + mu_L1 + mu_L2 + delta_L*L)*L
  dP <- rho_L*L - (rho_P + mu_P1 + mu_P2)*P
  dAh <- rho_P*P + phi_Ao*Ao - (rho_Ah + mu_A + mu_Ah2)*Ah
  dAr <- rho_Ah*Ah - (rho_Ar + mu_A +mu_Ar2)*Ar
  dAo <- rho_Ar*Ar - (phi_Ao + mu_A + mu_Ao2)*Ao
  #return(list(c(dE, dL, dP, dAh, dAr, dAo, temp, R)))
  return(list(c(dE, dL, dP, dAh, dAr, dAo)))
}

# Model parameters (All parameter values are based on Lutambi et al, 2013) and adapted to Anopheles species
b  = 200      # number of eggs laid per oviposition (range  100 - 400 eggs at a time) https://www.canada.ca/en/health-canada/services/pest-control-tips/mosquitoes.html
mu_E2 = 0.005    # extra natural mortality of eggs that is not dependent on environmental conditions
mu_L2 = 0.005 # Natural mortality of larvae
delta_L = 0.01 # For every 100 larvae, 1 larvae will probably die due to competition for resources (Assumed)
mu_P2 = 0.005 # Natural pupal mortality
rho_Ah =  2 # 0.46 # 2 (According to Cailly et al, 2012, this parameter takes the value of 2 day^-1) 
mu_Ah2 = 0.01 
rho_Ar =  1/2 # Takes 2 days for egg to develop and seek oviposition site
mu_Ar2 = 0.005 
mu_Ao2 = 0.01
K = 10^13  # From Okuneye et al, 

param <- c(b, 
           mu_E2, 
           mu_L2, 
           delta_L, 
           mu_P2, 
           rho_Ah, 
           mu_Ah2, 
           rho_Ar, 
           mu_Ar2, 
           mu_Ao2, 
           K)

# Initial conditions (I am not sure what values to use as initial conditions but I guess it doesn't matter much)
#y.initial <- c(100.0, 100.0, 100.0, 30.0, 30.0, 40.0, 6.4314, 41.0054)
y.initial <- c(100.0, 100.0, 100.0, 30.0, 30.0, 40.0)

# Time steps (Period with which to run the model)

start <- 0
end <- 365*5
times <- seq(start, end, by = 1)


# Numerical integration using the ode() function from deSolve package
out2 <-  ode(y = y.initial, func = pop.model2, times = times, parms = param, method = "lsoda") %>% 
  data.frame()

# Rename variable
# out2 <- out2 %>%
#   rename(eggs = X1, larvae = X2, pupae = X3, host.seeking.adults = X4,
#          resting.adults = X5, ovipositn.adults = X6, temp=X7, rain=X8)

out2 <- out2 %>%
  rename(eggs = X1, larvae = X2, pupae = X3, host.seeking.adults = X4,
         resting.adults = X5, ovipositn.adults = X6)

#Add a column which contains the total number of adults (A = Ah + Ar + Ao) to the data.frame
# out2 <- out2 %>% 
#   mutate(fixed.total.adults =  out2$host.seeking.adults + out2$resting.adults + out2$ovipositn.adults)
# # 

out2 <- out2 %>%
  mutate(vary.total.adults =  out2$host.seeking.adults + out2$resting.adults + out2$ovipositn.adults)


# Temp and rain need to be differentiated as it gets integrated through the function
# out2 <- out2 %>% 
#   mutate(temperature =  c(out2$temp[1],diff(out2$temp)*(1/1)), 
#          rainfall = c(out2$rain[1],diff(out2$rain)*(1/1)))

# This is when time step is 0.1. I changed it to 1 so we don't need this part.
#out2 <- out2 %>% 
#   mutate(temperature =  c(out2$temp[1],diff(out2$temp)*(1/0.1)), 
#         rainfall = c(out2$rain[1],diff(out2$rain)*(1/0.1)))

# I wanna drop temp and rain column
# df <- out2 %>%
#   subset(select = -c(temp, rain) )

#data <- write_csv(df, "data.csv")

df <- out2
# # Define a function to produce the plots
visualize <- function(p){
  # Set plot environment
  par(mfrow = c(3, 2), mar=c(2,4,2,2))    # 2 rows, 2 columns
  
  # Create plots
  plot(df$time, df$eggs, type = "l", xlab = "Time (days)", ylab = "Egg abundance",
       pch = 19, col = "blue")  # plot no. 1
  
  plot(df$time, df$larvae, type = "l", xlab = "Time (days)", ylab = "Larval abundance",
       pch = 19, col = "green")     # plot no. 2
  
  plot(df$time, df$pupae, type = "l", xlab = "Time (days)", ylab = "Pupal abundance",
       pch = 19, col = "yellow")     # plot no. 3
  
  plot(df$time, df$host.seeking.adults, type = "l", xlab = "Time (days)", ylab = "Adults seeking hosts",
       pch = 19, col = "red")    # plot no. 4
  
  plot(df$time, df$resting.adults, type = "l", xlab = "Time (days)", ylab = "Adults resting",
       pch = 19, col = "violet")    # plot no. 5
  
  plot(df$time, df$ovipositn.adults, type = "l", xlab = "Time (days)", ylab = "Adults ovipositing",
       pch = 19, col = "orange")    # plot no. 6
}

visualize(p)

#fixed.condt_sa <- df 
variable.condt_sa <- df 

#write_csv(fixed.condt_sa, file = "/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_sa.csv")
#write_csv(variable.condt_sa, file = "/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")
```


```{r}

##- --------------------------------------------------------------------------------------------------
#.  Figures for SA
##- --------------------------------------------------------------------------------------------------

# Load data
df_fix.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_sa.csv")
df_vary.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")

df.fix.sa <- df_fix.sa %>% 
  select(time, fixed.total.adults)

df.vary.sa <- df_vary.sa %>% 
  select(time, vary.total.adults)


df.sa.combined <- data.frame(time = df.fix.sa$time, constant = df.fix.sa$fixed.total.adults,
                                seasonal = df.vary.sa$vary.total.adults)

df.sa.combined <- df.sa.combined[1460:nrow(df.sa.combined), ]

df.sa.combined <- df.sa.combined %>% 
  pivot_longer(cols=c("constant", "seasonal"),
               names_to='Environment',
               values_to='value')

# Plot data 
population.sa <- ggplot(data = df.sa.combined, aes(x=time, y=value, col = Environment))+
  geom_line(linewidth = 1)+
  xlab("Time (days)") +
  ylab("Population") +
  theme_classic()

population.sa

# rename columns
df.sa.combined2 <- df.sa.combined %>% 
  rename(
    x = time,
    y = value
  )

# Change time to months
breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df.sa.combined2$y) * 1.1  # Adjust the buffer as needed (e.g., 1.1 for a 10% buffer)

# Define the colors you want to use
line_colors <- c("red", "darkgreen")

plot.sa <- ggplot(df.sa.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "South Africa",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.8)) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit)) +
scale_color_manual(values = line_colors)

plot.sa

# Save plot
ggsave(plot.sa, filename = "constant_and_seasonal_env_sa.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```

```{r}
data_seasonal_sa <- df.sa.combined2 %>% 
  filter(Environment == "seasonal")

seasonal_pop_abudance_sa <- sum(data_seasonal_sa$y)
seasonal_pop_abudance_sa
seasonal_pop_abudance_gh

seasonal_pop_abudance_gh/seasonal_pop_abudance_sa
```



I will put the figures together and label them to be included in the paper as a single figure
```{r}
library(cowplot)
library(magick)

# Load saved figures
plot_algeria <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/constant_and_seasonal_env_algeria.png", scale = 1)
plot_ghana <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/constant_and_seasonal_env_ghana.png", scale = 1)
plot_kenya <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/constant_and_seasonal_env_kenya.png", scale = 1)
plot_sa <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/constant_and_seasonal_env_sa.png", scale = 1)

# Numbering the plots
numbered_plot_algeria <- plot_algeria + draw_label("(a)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14) # Here the font normal
numbered_plot_sa <- plot_sa + draw_label("(b)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14)
numbered_plot_kenya <- plot_kenya + draw_label("(c)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14)
numbered_plot_ghana <- plot_ghana + draw_label("(d)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14)

# Combine the plots
combined_plot <- plot_grid(
  numbered_plot_algeria, 
  numbered_plot_sa, 
  numbered_plot_kenya, 
  numbered_plot_ghana,
  ncol = 2
)

# Display the combined plot
print(combined_plot)

#ggsave(combined_plot, filename = "constant_vrs_seasonal_combined.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```



## I am doing the same as above in one big chunk

```{r}
# Clear space
rm(list = ls())

# Load data (on macbook)
df_fix <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_Alg.csv")
df_vary <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_Alg.csv")

# Load data (on mac-mini)
#df_fix <- read.csv("/Users/josephbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_Alg.csv")
#df_vary <- read.csv("/Users/josephbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_Alg.csv")

# Select needed columns
df.fix <- df_fix %>% 
  select(time, fixed.total.adults)

# Select needed columns
df.vary <- df_vary %>% 
  select(time, vary.total.adults)

# Put together into one dataframe
df.alg.combined <- data.frame(time = df.fix$time, constant = df.fix$fixed.total.adults,
                              seasonal = df.vary$vary.total.adults)

# Select the year 5 data 
df.alg.combined <- df.alg.combined[1461:nrow(df.alg.combined), ]

# Convert data to the long format
df.alg.combined <- df.alg.combined %>% 
  pivot_longer(cols=c("constant", "seasonal"),
               names_to='Environment',
               values_to='value')

# rename columns
df.alg.combined2 <- df.alg.combined %>% 
  rename(
    x = time,
    y = value
  )

# Change time to months
breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df.alg.combined2$y) * 1.1  # Adjust the buffer as needed (e.g., 1.1 for a 10% buffer)


plot.alg <- ggplot(df.alg.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = paste("(a)", "Algeria", sep = " "),  # Add a space between (a) and Algeria
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.8),  # Adjust the values to move the legend
    legend.box = "horizontal",  # Use "vertical" for a vertical box
    legend.box.background = element_rect(color = "black", size = 1)  # Rectangle around the legend
  ) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit))

plot.alg

##- --------------------------------------------------------------------------------------------------
#. I use saved data from the above simulation analysis to produce figures for Ghana
##- --------------------------------------------------------------------------------------------------

# Load data (on macbook)
df_fix.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_gh.csv")
df_vary.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")

# Load data (on macmini)
# df_fix.gh <- read.csv("/Users/josephbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_gh.csv")
# df_vary.gh <- read.csv("/Users/josephbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")

df.fix.gh <- df_fix.gh %>% 
  select(time, fixed.total.adults)

df.vary.gh <- df_vary.gh %>% 
  select(time, vary.total.adults)


df.gh.combined <- data.frame(time = df.fix.gh$time, constant = df.fix.gh$fixed.total.adults,
                             seasonal = df.vary.gh$vary.total.adults)

df.gh.combined <- df.gh.combined[1460:nrow(df.gh.combined), ]

df.gh.combined <- df.gh.combined %>% 
  pivot_longer(cols=c("constant", "seasonal"),
               names_to='Environment',
               values_to='value')

# rename columns
df.gh.combined2 <- df.gh.combined %>% 
  rename(
    x = time,
    y = value
  )

# Change time to months
breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df.gh.combined2$y) * 1.1  # Adjust the buffer as needed (e.g., 1.1 for a 10% buffer)

plot.gh <- ggplot(df.gh.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "(d) Ghana",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.7),  # Adjust the values to move the legend
    legend.box = "horizontal",  # Use "vertical" for a vertical box
    legend.box.background = element_rect(color = "black", size = 1)  # Rectangle around the legend
  ) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit))

plot.gh

##- --------------------------------------------------------------------------------------------------
#.  Figures for Kenya
##- --------------------------------------------------------------------------------------------------

# Load data
df_fix.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_kenya.csv")
df_vary.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")

df.fix.kenya <- df_fix.kenya %>% 
  select(time, fixed.total.adults)

df.vary.kenya <- df_vary.kenya %>% 
  select(time, vary.total.adults)


df.kenya.combined <- data.frame(time = df.fix.kenya$time, constant = df.fix.kenya$fixed.total.adults,
                                seasonal = df.vary.kenya$vary.total.adults)

df.kenya.combined <- df.kenya.combined[1460:nrow(df.kenya.combined), ]

df.kenya.combined <- df.kenya.combined %>% 
  pivot_longer(cols=c("constant", "seasonal"),
               names_to='Environment',
               values_to='value')

# rename columns
df.kenya.combined2 <- df.kenya.combined %>% 
  rename(
    x = time,
    y = value
  )

# Change time to months
breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df.kenya.combined2$y) * 1.1  # Adjust the buffer as needed (e.g., 1.1 for a 10% buffer)

plot.kenya <- ggplot(df.kenya.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "(c)  Kenya",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.4),  # Adjust the values to move the legend
    legend.box = "horizontal",  # Use "vertical" for a vertical box
    legend.box.background = element_rect(color = "black", size = 1)  # Rectangle around the legend
  ) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit))

plot.kenya


##- --------------------------------------------------------------------------------------------------
#.  Figures for SA
##- --------------------------------------------------------------------------------------------------

# Load data
df_fix.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_sa.csv")
df_vary.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")

df.fix.sa <- df_fix.sa %>% 
  select(time, fixed.total.adults)

df.vary.sa <- df_vary.sa %>% 
  select(time, vary.total.adults)


df.sa.combined <- data.frame(time = df.fix.sa$time, constant = df.fix.sa$fixed.total.adults,
                             seasonal = df.vary.sa$vary.total.adults)

df.sa.combined <- df.sa.combined[1460:nrow(df.sa.combined), ]

df.sa.combined <- df.sa.combined %>% 
  pivot_longer(cols=c("constant", "seasonal"),
               names_to='Environment',
               values_to='value')

# rename columns
df.sa.combined2 <- df.sa.combined %>% 
  rename(
    x = time,
    y = value
  )

# Change time to months
breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df.sa.combined2$y) * 1.1  # Adjust the buffer as needed (e.g., 1.1 for a 10% buffer)

plot.sa <- ggplot(df.sa.combined2, aes(x, y, col = Environment)) +
  geom_line(lwd = 1) +
  labs(title = "(b)  South Africa",
       x = "",
       y = "Mosquito abundance") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12),
    legend.position = c(0.1, 0.8),  # Adjust the values to move the legend
    legend.box = "horizontal",  # Use "vertical" for a vertical box
    legend.box.background = element_rect(color = "black", size = 1)  # Rectangle around the legend
  ) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit))

plot.sa


# Install and load the 'cowplot' package
library(cowplot)

# Adjust legend size and position for each plot
plot.alg1 <- plot.alg + theme(legend.position = c(0.25, 0.6))
plot.sa1 <- plot.sa + theme(legend.position = "None")
plot.kenya1 <- plot.kenya + theme(legend.position = "None")
plot.gh1 <- plot.gh + theme(legend.position = "None")

library(patchwork)
# Combine the plots using the patchwork package
combined_plots1 <- (plot.alg1 | plot.sa1) / (plot.kenya1 | plot.gh1)

# Print the combined plots
print(combined_plots1)


# Save plot
#ggsave(combined_plots1, filename = "constant_and_seasonal_env_combined.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")



```


## Comparing regional differences in adult population dynamics 

The graph is plotted from data that was generated after running the model for 5 years. The 5th year generates this plot. We do this to allow initial conditions to neutralize in the system and the system going to kind of equilibrium. 

- Think about why this Fig. is important for the research objectives.


```{r}
# Regional population dynamics of adult mosquitoes

# Load package that helps to load data
library(rio)

# Adult abundance data for different regions
df.alg <- import("/Users/jbaafi/Documents/Github/population-model/data/df_algeria.csv")
df.gh <- import("/Users/jbaafi/Documents/Github/population-model/data/df_ghana.csv")
df.kenya <- import("/Users/jbaafi/Documents/Github/population-model/data/df_kenya.csv")
df.sa <- import("/Users/jbaafi/Documents/Github/population-model/data/df_SA.csv")

# Another saved data (used in the next chunk)
 #df.alg <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_alg.csv")
# df.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_gh.csv")
# df.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_kenya.csv")
# df.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/fixed.condt_sa.csv")


df2 <- data.frame(time = df.alg$time, Algeria=df.alg$total.adults, Ghana=df.gh$total.adults, Kenya=df.kenya$total.adults, 
                  South.Africa=df.sa$total.adults)

df2 <- df2[1460:nrow(df2), ]

df2 <- df2 %>% 
  pivot_longer(cols=c("Algeria", "Ghana", "Kenya", "South.Africa"),
               names_to='Region',
               values_to='population')

ggp <- ggplot(df2, aes(x=time, y=population, col = Region))+
  geom_line()+
  xlab("Time (days)") +
  ylab("Female adult abundance") +
  ggtitle("Number of female adult mosquitoes in different regions") +
  theme_bw()

ggp +                               # Set origin of axes to zero
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))

# At this point, I will change time on x-axis from days to months

df2 <- df2 %>% 
  rename(
    x = time,
    y = population
  )

breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df2$y) * 1.1

adult <- ggplot(df2, aes(x, y, col = Region)) +
  geom_line(lwd = 1) +
  labs(title = "",
       x = "",
       y = "Number of adults") +
  theme_classic() +
  theme(axis.text.x = element_text(colour = "grey20", size = 10, angle = 60,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        text = element_text(size = 10)) +
  scale_x_continuous(breaks=breakfunc, labels=labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit))

adult

# Save the plot

#ggsave(adult, filename = "adult_abundance.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")
```


```{r}
# Regional population dynamics of adult mosquitoes

# Load package that helps to load data
library(rio)

# Another saved data
df.alg <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_alg.csv")
df.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")
df.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")
df.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")


df2 <- data.frame(time = df.alg$time, Algeria=df.alg$vary.total.adults, Ghana=df.gh$vary.total.adults, Kenya=df.kenya$vary.total.adults, South.Africa=df.sa$vary.total.adults)

df2 <- df2[1461:nrow(df2), ]

df2 <- df2 %>% 
  pivot_longer(cols=c("Algeria", "Ghana", "Kenya", "South.Africa"),
               names_to='Region',
               values_to='population')

ggp <- ggplot(df2, aes(x=time, y=population, col = Region))+
  geom_line()+
  xlab("Time (days)") +
  ylab("Female adult abundance") +
  ggtitle("Number of female adult mosquitoes in different regions") +
  theme_bw()

ggp +                               # Set origin of axes to zero
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))

# At this point, I will change time on x-axis from days to months

df2 <- df2 %>% 
  rename(
    x = time,
    y = population
  )

breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

upper_limit <- max(df2$y) * 1.1

# Define the colors you want to use for each region
line_colors <- c("purple", "darkred", "orange", "darkgreen")

# Your existing plot code
adult <- ggplot(df2, aes(x, y, col = Region)) +
  geom_line(lwd = 1) +
  labs(title = "",
       x = "",
       y = "Number of adults") +
  theme_classic() +
  theme(axis.text.x = element_text(colour = "grey20", size = 11, angle = 60,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 11),
        text = element_text(size = 12)) +
  scale_x_continuous(breaks = breakfunc, labels = labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, upper_limit)) +
  scale_color_manual(values = line_colors)

# Save the plot
#ggsave(adult, filename = "adult_abundance.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")
```




## Percentage of yearly adult total for each month.

This is a 5 year average. 

```{r}
# Ain Mahbel, Algeria
df.alg <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_alg.csv")

# Calculate the total count of adult mosquitoes for the entire year
yearly_total.alg <- sum(df.alg$vary.total.adults)

# Create a column called count to data which is same as total adult numbers
df.alg <- df.alg %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
mosquito_data <- df.alg %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a bar plot displaying the monthly percentages:
ggplot(mosquito_data, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "red") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("Month") +
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes") +
  theme_bw()



# Modify the plot
ggplot(mosquito_data, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "red") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +  # Change size of y-axis label
  xlab("Month") +  # Change size of x-axis label
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 11),   # Change size of x-axis ticks
        axis.text.y = element_text(size = 11),   # Change size of y-axis ticks
        axis.title.x = element_text(size = 12),  # Change size of x-axis label
        axis.title.y = element_text(size = 12))  # Change size of y-axis label

```


```{r}
# Same as above for Ghana
df.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")

df.gh <- df.gh %>% 
  mutate(count = vary.total.adults)

# Extract the month and year information from the time column. Then, calculate the monthly percentages considering data from all five years:
mosquito_data.gh <- df.gh %>%
  mutate(year = as.integer(format(as.Date(time, origin = "2021-01-01"), "%Y")),
         month = as.integer(format(as.Date(time, origin = "2021-01-01"), "%m"))) %>%
  group_by(year, month) %>%
  summarise(monthly_total = sum(count)) %>%
  group_by(month) %>%
  summarise(yearly_total = sum(monthly_total)) %>%
  mutate(monthly_percentage = (yearly_total / sum(yearly_total)) * 100)

# Show the plot
ggplot(mosquito_data.gh, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("Month") +
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes") +
  theme_bw()

```



```{r}
# Same as above for Kenya
df.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")

df.kenya <- df.kenya %>% 
  mutate(count = vary.total.adults)

# Extract the month and year information from the time column. Then, calculate the monthly percentages considering data from all five years:
mosquito_data.kenya <- df.kenya %>%
  mutate(year = as.integer(format(as.Date(time, origin = "2021-01-01"), "%Y")),
         month = as.integer(format(as.Date(time, origin = "2021-01-01"), "%m"))) %>%
  group_by(year, month) %>%
  summarise(monthly_total = sum(count)) %>%
  group_by(month) %>%
  summarise(yearly_total = sum(monthly_total)) %>%
  mutate(monthly_percentage = (yearly_total / sum(yearly_total)) * 100)

# Show the plot
ggplot(mosquito_data.kenya, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("Month") +
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes")+
  theme_bw()

```


```{r}
# Same as above for SA
df.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")

df.sa <- df.sa %>% 
  mutate(count = vary.total.adults)

# Extract the month and year information from the time column. Then, calculate the monthly percentages considering data from all five years:
mosquito_data.sa <- df.sa %>%
  mutate(year = as.integer(format(as.Date(time, origin = "2021-01-01"), "%Y")),
         month = as.integer(format(as.Date(time, origin = "2021-01-01"), "%m"))) %>%
  group_by(year, month) %>%
  summarise(monthly_total = sum(count)) %>%
  group_by(month) %>%
  summarise(yearly_total = sum(monthly_total)) %>%
  mutate(monthly_percentage = (yearly_total / sum(yearly_total)) * 100)

# Show the plot
ggplot(mosquito_data.sa, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "purple") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("Month") +
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes") +
  theme_bw()

```

## Combine plots into single plot

**1. Understanding the Y-Axis (Percentage of Yearly Total):
Y-Axis Values: The y-axis represents the percentage of the total yearly adult mosquito population that each month contributes. Higher values indicate months with a larger proportion of mosquitoes, whereas lower values indicate months with a smaller proportion.
2. Understanding the X-Axis (Months):
Months: The x-axis represents the months of the year (from January to December). Each month's bar shows the percentage of mosquitoes for that specific month across all the years of data.
3. Comparing Monthly Patterns Across Regions:
Peak Months: Look for months with the tallest bars in each region. These months represent the periods when the mosquito population is at its peak in each region. Understanding which months have the highest percentages can indicate the months with the highest mosquito activity.

Low Months: Conversely, months with shorter bars indicate periods when the mosquito population is relatively low. This could be due to factors such as colder weather or lower rainfall, which are less conducive to mosquito breeding.

4. Comparing Seasonality Across Regions:
Comparing Peaks: Compare the peak months across different regions. Do these peak months align? If not, it suggests variations in the seasonality of mosquito populations among regions.

Comparing Patterns: Look for similarities or differences in the patterns between regions. For instance, if two regions show a peak in mosquito activity during the same months, it might indicate a common environmental trigger, such as rainy seasons or specific temperatures.

5. Implications for Public Health and Vector Control:
Disease Transmission: Understanding when mosquito populations are at their peak can help anticipate periods of higher disease transmission risk. It allows health authorities to implement targeted prevention and control measures, such as mosquito net distribution or insecticide spraying, during these high-risk months.

Vector Control: Vector control programs can use this information to optimize their efforts. For example, during peak months, they might intensify mosquito surveillance, implement larval source reduction, or conduct targeted insecticide spraying to reduce the mosquito population.

6. Environmental Insights:
Climate Impact: Changes in mosquito patterns can be influenced by climate change. Comparing the plots over multiple years might provide insights into the impact of climate variability on mosquito populations.
7. Limitations and Further Analysis:
Data Quality: Consider the quality and consistency of the data. Sudden spikes or drops in percentages should be investigated for potential data anomalies or errors.

Further Analysis: For a more comprehensive analysis, consider additional factors such as temperature, rainfall, or humidity data, which can help explain the observed patterns and provide a more accurate understanding of mosquito behavior.

Interpreting these plots in the context of local climate patterns and disease epidemiology can provide valuable information for public health strategies, especially in regions where mosquito-borne diseases are a concern.


```{r}

# Assume you have separate data frames for each region: algeria_data, ghana_data, kenya_data, sa_data

# Creating separate plots for each region
plot_algeria <- ggplot(mosquito_data, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("Month") +
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes in Algeria") 

plot_ghana <- ggplot(mosquito_data.gh, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("Month") +
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes in Ghana")

plot_kenya <- ggplot(mosquito_data.kenya, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("Month") +
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes in Kenya")

plot_sa <- ggplot(mosquito_data.sa, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("Month") +
  ggtitle("Monthly Percentage of Yearly Total Adult Mosquitoes in South Africa")

# Arranging the plots in a grid using facet_wrap
combined_plot <- cowplot::plot_grid(plot_algeria, plot_ghana, plot_kenya, plot_sa, ncol = 2)

# Print the combined plot
print(combined_plot)

```


```{r}
# Same as previous chunk with a little modification to the code
# Creating separate plots for each region
plot_algeria <- ggplot(mosquito_data, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "purple") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("") +
  ggtitle("Algeria") +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 1),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12))

plot_ghana <- ggplot(mosquito_data.gh, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "darkred") +
  scale_x_discrete(labels = month.abb) +
  ylab("") +
  xlab("") +
  ggtitle("Ghana")  +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 1),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12))

plot_kenya <- ggplot(mosquito_data.kenya, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "orange") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("") +
  ggtitle("Kenya")  +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 1),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12))

plot_sa <- ggplot(mosquito_data.sa, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_x_discrete(labels = month.abb) +
  ylab("") +
  xlab("") +
  ggtitle("South Africa") +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 11, angle = 60, hjust = 1),
    axis.text.y = element_text(colour = "grey20", size = 11),
    text = element_text(size = 12))

# Arranging the plots in a grid using facet_wrap
combined_plot <- cowplot::plot_grid(plot_algeria, plot_sa, plot_kenya, plot_ghana, ncol = 2)

# Print the combined plot
print(combined_plot)

#ggsave(combined_plot, filename = "percentage_of_yearly_total.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```


```{r, warning=FALSE, message=FALSE, fig.height=15, fig.width=25}
# Same as previous chunk with a little modification to the code
# Creating separate plots for each region
plot_algeria <- ggplot(mosquito_data, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "purple") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("") +
  ggtitle("Algeria") +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 18, angle = 60, hjust = 1),
    axis.text.y = element_text(colour = "grey20", size = 18),
    text = element_text(size = 22))

plot_ghana <- ggplot(mosquito_data.gh, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "darkred") +
  scale_x_discrete(labels = month.abb) +
  ylab("") +
  xlab("") +
  ggtitle("Ghana")  +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 18, angle = 60, hjust = 1),
    axis.text.y = element_text(colour = "grey20", size = 18),
    text = element_text(size = 22))

plot_kenya <- ggplot(mosquito_data.kenya, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "orange") +
  scale_x_discrete(labels = month.abb) +
  ylab("Percentage of Yearly Total") +
  xlab("") +
  ggtitle("Kenya")  +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 18, angle = 60, hjust = 1),
    axis.text.y = element_text(colour = "grey20", size = 18),
    text = element_text(size = 22))

plot_sa <- ggplot(mosquito_data.sa, aes(x = factor(month), y = monthly_percentage)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_x_discrete(labels = month.abb) +
  ylab("") +
  xlab("") +
  ggtitle("South Africa") +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 18, angle = 60, hjust = 1),
    axis.text.y = element_text(colour = "grey20", size = 18),
    text = element_text(size = 22))

# Arranging the plots in a grid using facet_wrap
combined_plot <- cowplot::plot_grid(plot_algeria, plot_sa, plot_kenya, plot_ghana, ncol = 2)

# Print the combined plot
print(combined_plot)

#ggsave(combined_plot, filename = "percentage_of_yearly_total.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```



## Plot to investigate whether adult mosquito control needs to be concentrated across a few months or should be spread out over a year. 


This is the chunk that produces the required figure for Algeria

```{r}

# Ain Mahbel, Algeria
df.alg <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_alg.csv")

# Calculate the total count of adult mosquitoes for the entire year
#yearly_total.alg <- sum(df.alg$vary.total.adults)

# Create a column called count to data which is same as total adult numbers
df.alg <- df.alg %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
algeria_data <- df.alg %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
algeria_data <- algeria_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
algeria_sorted_data <- algeria_data[order(-algeria_data$monthly_percentage),]

# Save the reordered data as a new data frame
new_data <- data.frame(months = factor(algeria_sorted_data$months),
                       monthly_percentage = algeria_sorted_data$monthly_percentage)


algeria_data <- new_data %>% 
  mutate(cum_fraction = cumsum(monthly_percentage))


# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
ggplot(algeria_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("Fraction of adults (%)") +
  xlab("Month") +
  ggtitle("Cumulative sum of adult mosquitoes - Algeria") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50



# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
algeria_step_plot <- ggplot(algeria_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("Percentage of adults (%)") +
  xlab("Month") +
  ggtitle("Cumulative sum of adult mosquitoes - Algeria") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(algeria_data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

algeria_step_plot
```
From the figure above, it is clear that if you focus control interventions in the months of August, September and October, you control about 80% of the mosquito population. The months of November to April require mostly no interventions at all. This insight is really helpful for vector-borne disease control and optimal use of resources in this region. 


This chunk produces the figure for Ghana
```{r}
df.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")

# Create a column called count to data which is same as total adult numbers
df.gh <- df.gh %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
ghana_data <- df.gh %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
ghana_data <- ghana_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
ghana_sorted_data <- ghana_data[order(-ghana_data$monthly_percentage),]

# Save the reordered data as a new data frame
new_data <- data.frame(months = factor(ghana_sorted_data$months),
                       monthly_percentage = ghana_sorted_data$monthly_percentage)


ghana_data <- new_data %>% 
  mutate(cum_fraction = cumsum(monthly_percentage))


# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
ggplot(ghana_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", linewidth = 1) +
  ylab("Fraction of adults (%)") +
  xlab("Month") +
  ggtitle("Cumulative sum of adult mosquitoes - Ghana") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50



# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
ghana_step_plot <- ggplot(ghana_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("Percentage of adults (%)") +
  xlab("Month") +
  ggtitle("Cumulative sum of adult mosquitoes - Ghana") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(ghana_data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

ghana_step_plot
```

It looks like controls should be focused all months across the year since we have a significant fraction of mosquitoes across the year.


This chunk produces the figure for Kenya
```{r}
df.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")

# Create a column called count to data which is same as total adult numbers
df.kenya <- df.kenya %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
kenya_data <- df.kenya %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
kenya_data <- kenya_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
kenya_sorted_data <- kenya_data[order(-kenya_data$monthly_percentage),]

# Save the reordered data as a new data frame
new_data <- data.frame(months = factor(kenya_sorted_data$months),
                       monthly_percentage = kenya_sorted_data$monthly_percentage)


kenya_data <- new_data %>% 
  mutate(cum_fraction = cumsum(monthly_percentage))


# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
ggplot(kenya_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", linewidth = 1) +
  ylab("Fraction of adults (%)") +
  xlab("Month") +
  ggtitle("Cumulative sum of adult mosquitoes - Kenya") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50



# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
kenya_step_plot <- ggplot(kenya_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("Percentage of adults (%)") +
  xlab("Month") +
  ggtitle("Cumulative sum of adult mosquitoes - Kenya") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(kenya_data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

kenya_step_plot
```
This also looks like we need to implement controls in each month across the year to keep mosquitoes numbers low and reduce the risk of disease spread. 


This chunk produces the figure for SA
```{r}
df.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")

# Create a column called count to data which is same as total adult numbers
df.sa <- df.sa %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
sa_data <- df.sa %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
sa_data <- sa_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
sa_sorted_data <- sa_data[order(-sa_data$monthly_percentage),]

# Save the reordered data as a new data frame
new_data <- data.frame(months = factor(sa_sorted_data$months),
                       monthly_percentage = sa_sorted_data$monthly_percentage)


sa_data <- new_data %>% 
  mutate(cum_fraction = cumsum(monthly_percentage))


# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
ggplot(sa_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", linewidth = 1) +
  ylab("Fraction of adults (%)") +
  xlab("Month") +
  ggtitle("Cumulative sum of adult mosquitoes - SA") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50



# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
sa_step_plot <- ggplot(sa_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("Percentage of adults (%)") +
  xlab("Month") +
  ggtitle("Cumulative sum of adult mosquitoes - SA") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(sa_data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

sa_step_plot
```


Combine plots of cummulative sum of adult mosquitoes for different regions.

```{r}

# Creating separate plots for each region
algeria_step_plot <- ggplot(algeria_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("Percentage of adults (%)") +
  xlab("") +
  ggtitle("Algeria") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(algeria_data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

ghana_step_plot <- ggplot(ghana_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("") +
  xlab("") +
  ggtitle("Ghana") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(ghana_data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

kenya_step_plot <- ggplot(kenya_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("Percentage of adults (%)") +
  xlab("Month") +
  ggtitle("Kenya") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(kenya_data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

sa_step_plot <- ggplot(sa_data, aes(x = months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("") +
  xlab("Month") +
  ggtitle("SA") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(sa_data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

# Arranging the plots in a grid using facet_wrap
combined_step_plots <- cowplot::plot_grid(algeria_step_plot, ghana_step_plot, kenya_step_plot, sa_step_plot, ncol = 2)

# Print the combined plot
combined_step_plots

#ggsave(combined_step_plots, filename = "step_plots.png", path = "/Users/jbaafi/Desktop/plots")

```


Now I want to put all the graphs into one plot. I am trying out code in the following chunks, no serious chunk yet. 

```{r}

library(ggtext)

# Create a data frame with months and counts
months_list <- month.abb
count <- c(200, 250, 500, 100, 75, 80, 125, 100, 244, 400, 120, 98)

numeric_months <- 1:12

total_count <- sum(count) 

# Calculate the fraction (percentage) of count for each month
fraction <- count / total_count * 100

# Create a new data frame with months, count, and fraction
data <- data.frame(months = factor(months_list, levels = months_list[order(fraction, decreasing = TRUE)]), 
                   numeric_months = numeric_months,
                   count = count,
                   fraction = fraction)
# Reorder the data frame based on decreasing fraction
sorted_data <- data[order(-data$fraction),]

# Save the reordered data as a new data frame
new_data <- data.frame(months = factor(sorted_data$months),
                       ranked_months = factor(1:12),
                       fraction = sorted_data$fraction)


data <- new_data %>% 
  mutate(cum_fraction = cumsum(fraction))


# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
ggplot(data, aes(x = ranked_months, y = cum_fraction, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("") +
  xlab("Month") +
  ggtitle("SA") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(data$cum_fraction)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

```


```{r}
# Create a data frame with months and counts
months_list <- month.abb
count <- c(2, 5, 100, 80, 95, 100, 150, 120, 10, 4, 12, 2)

numeric_months <- 1:12

total_count <- sum(count) 

# Calculate the fraction (percentage) of count for each month
fraction <- count / total_count * 100

# Create a new data frame with months, count, and fraction
data2 <- data.frame(months = factor(months_list, levels = months_list[order(fraction, decreasing = TRUE)]), 
                   numeric_months = numeric_months,
                   count = count,
                   fraction = fraction)
# Reorder the data frame based on decreasing fraction
sorted_data2 <- data2[order(-data2$fraction),]

# Save the reordered data as a new data frame
new_data2 <- data.frame(months2 = factor(sorted_data2$months),
                       ranked_months = factor(1:12),
                       fraction2 = sorted_data2$fraction)


data2 <- new_data2 %>% 
  mutate(cum_fraction2 = cumsum(fraction2))


# Plotting the stair step function with months on the x-axis and fraction of count on the y-axis
ggplot(data2, aes(x = ranked_months, y = cum_fraction2, group = 1)) +
  geom_step(color = "skyblue", size = 1) +
  ylab("") +
  xlab("Month") +
  ggtitle("SA") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  ylim(0, max(data2$cum_fraction2)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50

```

I combine the previous two data and provide a combined plot. 

```{r}
#data
#data2

data_merged <- inner_join(data, data2) 

# Plotting the stair step function with months on the x-axis and cum_fraction/cum_fraction2 on the y-axis
ggplot(data_merged, aes(x = ranked_months, group = 1)) +
  geom_step(aes(y = cum_fraction), color = "skyblue", size = 1) +
  geom_step(aes(y = cum_fraction2), color = "lightgreen", size = 1) +
  ylab("") +
  xlab("Month") +
  ggtitle("SA") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1)  # Add horizontal dashed line at y=50


# Assuming cum_fraction and cum_fraction2 are columns in data_merged
ggplot(data_merged, aes(x = ranked_months, group = 1)) +
  geom_step(aes(y = cum_fraction, color = "Cum Fraction 1"), size = 1) +
  geom_step(aes(y = cum_fraction2, color = "Cum Fraction 2"), size = 1) +
  ylab("") +
  xlab("Month") +
  ggtitle("SA") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better visibility
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1) +  # Add horizontal dashed line at y=50
  scale_color_manual(values = c("Cum Fraction 1" = "skyblue", "Cum Fraction 2" = "lightgreen"), 
                     name = "Legend Title")  # Define colors and legend title


ggplot(data_merged, aes(x = ranked_months, group = 1)) +
  geom_step(aes(y = cum_fraction, color = "Cum Fraction 1"), size = 1) +
  geom_step(aes(y = cum_fraction2, color = "Cum Fraction 2"), size = 1) +
  ylab("") +
  xlab("Month") +
  ggtitle("SA") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1)) +  # Rotate x-axis labels for better visibility
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1) +  # Add horizontal dashed line at y=50
  scale_color_manual(values = c("Cum Fraction 1" = "skyblue", "Cum Fraction 2" = "lightgreen"), 
                     name = "Legend Title") +  # Define colors and legend title
  annotate("text", x = max(as.numeric(as.character(data_merged$ranked_months))), 
           y = tail(data_merged$cum_fraction, 1), label = "Cum Fraction 1",
           vjust = 6, hjust = 1, color = "skyblue") +  # Add label for Cum Fraction 1
  annotate("text", x = max(as.numeric(as.character(data_merged$ranked_months))), 
           y = tail(data_merged$cum_fraction2, 1), label = "Cum Fraction 2",
           vjust = 0, hjust = 1, color = "lightgreen")  # Add label for Cum Fraction 2

```

Now I will use actual data to provide a plot similar to the ones above. 

```{r}
df.alg <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_alg.csv")
df.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")
df.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")
df.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")

# Calculate the total count of adult mosquitoes for the entire year
#yearly_total.alg <- sum(df.alg$vary.total.adults)

# Create a column called count to data which is same as total adult numbers
df.alg <- df.alg %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
algeria_data <- df.alg %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
algeria_data <- algeria_data %>%
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
algeria_sorted_data <- algeria_data[order(-algeria_data$monthly_percentage),]

# Save the reordered data as a new data frame
new_data <- data.frame(months_alg = factor(algeria_sorted_data$months),
                       ranked_months = factor(1:12),
                       monthly_percentage_alg = algeria_sorted_data$monthly_percentage,
                       month_sorted_alg = algeria_sorted_data$month)


algeria_data <- new_data %>% 
  mutate(cum_fraction_alg = cumsum(monthly_percentage_alg))


## ---------------------------------------- ----------------------------------------
## Ghana follows here
## ---------------------------------------- ----------------------------------------
 
# Create a column called count to data which is same as total adult numbers
df.gh <- df.gh %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
ghana_data <- df.gh %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
ghana_data <- ghana_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
ghana_sorted_data <- ghana_data[order(-ghana_data$monthly_percentage),]

# Save the reordered data as a new data frame
new_data <- data.frame(months_gh = factor(ghana_sorted_data$months),
                       ranked_months = factor(1:12),
                       monthly_percentage_gh = ghana_sorted_data$monthly_percentage)


ghana_data <- new_data %>% 
  mutate(cum_fraction_gh = cumsum(monthly_percentage_gh))

## ---------------------------------------- ----------------------------------------
## Kenya follows here
## ---------------------------------------- ----------------------------------------
# Create a column called count to data which is same as total adult numbers
df.kenya <- df.kenya %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
kenya_data <- df.kenya %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
kenya_data <- kenya_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
kenya_sorted_data <- kenya_data[order(-kenya_data$monthly_percentage),]

# Save the reordered data as a new data frame
new_data <- data.frame(months_kenya = factor(kenya_sorted_data$months),
                       ranked_months = factor(1:12),
                       monthly_percentage_kenya = kenya_sorted_data$monthly_percentage)


kenya_data <- new_data %>% 
  mutate(cum_fraction_kenya = cumsum(monthly_percentage_kenya))

## ---------------------------------------- ----------------------------------------
## SA section here
## ---------------------------------------- ----------------------------------------
# Create a column called count to data which is same as total adult numbers
df.sa <- df.sa %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
sa_data <- df.sa %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
sa_data <- sa_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
sa_sorted_data <- sa_data[order(-sa_data$monthly_percentage),]

# Save the reordered data as a new data frame
new_data <- data.frame(months_sa = factor(sa_sorted_data$months),
                       ranked_months = factor(1:12),
                       monthly_percentage_sa = sa_sorted_data$monthly_percentage)


sa_data <- new_data %>% 
  mutate(cum_fraction_sa = cumsum(monthly_percentage_sa))


## ---------------------------------------- ----------------------------------------
## Merge data and proceed to plot
## ---------------------------------------- ----------------------------------------
 
merged_data <- inner_join(algeria_data, ghana_data) %>% 
  inner_join(kenya_data) %>% 
  inner_join(sa_data)

# Plot
cum_plot <- ggplot(merged_data, aes(x = ranked_months, group = 1)) +
  geom_step(aes(y = cum_fraction_alg, color = "Algeria"), size = 1) +
  geom_step(aes(y = cum_fraction_gh, color = "Ghana"), size = 1) +
  geom_step(aes(y = cum_fraction_kenya, color = "Kenya"), size = 1) +
  geom_step(aes(y = cum_fraction_sa, color = "South Africa"), size = 1) +
  ylab("Cummulative fraction of mosquitoes (%)") +
  xlab("Months up to rank") +
  ggtitle("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1)) +  # Rotate x-axis labels for better visibility
   ylim(0, max(merged_data$cum_fraction_alg)) + # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1) +  # Add horizontal dashed line at y=50
  scale_color_manual(values = c("Algeria" = "yellow", "Ghana" = "red", "Kenya" = "orange", "South Africa" = "blue"), name = "Region") 

cum_plot

cummulative_plot <- ggplot(merged_data, aes(x = ranked_months, group = 1)) +
  geom_step(aes(y = cum_fraction_alg, color = "Algeria"), size = 1.2, color = "purple") +
  geom_step(aes(y = cum_fraction_gh, color = "Ghana"), size = 1.2, color = "darkred") +
  geom_step(aes(y = cum_fraction_kenya, color = "Kenya"), size = 1.2, color = "orange") +
  geom_step(aes(y = cum_fraction_sa, color = "South Africa"), size = 1.2, color = "darkgreen") +
  ylab("Cumulative fraction of mosquitoes (%)") +
  xlab("Months up to rank") +
  ggtitle("") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 1, size = 12),  # Adjust size as needed
    axis.text.y = element_text(size = 12),  # Adjust size as needed
    axis.title = element_text(size = 12),
    text = element_text(size = 12),
    legend.position = "none"  # Remove legend
  ) +
  ylim(0, max(merged_data$cum_fraction_alg)) +  # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1.2) +
  geom_text(data = merged_data, 
            aes(x = max(as.numeric(levels(ranked_months))), 
                y = tail(cum_fraction_alg, 1),
                label = "ALGR"),
            vjust = -0.5, hjust = 6, color = "purple") +
  geom_text(data = merged_data, 
            aes(x = max(as.numeric(levels(ranked_months))), 
                y = tail(cum_fraction_gh, 1),
                label = "GHAN"),
            vjust = 4, hjust = 5, color = "darkred") +
  geom_text(data = merged_data, 
            aes(x = max(as.numeric(levels(ranked_months))), 
                y = tail(cum_fraction_kenya, 1),
                label = "KENY"),
            vjust = 6, hjust = 2, color = "orange") +
  geom_text(data = merged_data, 
            aes(x = max(as.numeric(levels(ranked_months))), 
                y = tail(cum_fraction_sa, 1),
                label = "SA"),
            vjust = 5.4, hjust = 20, color = "darkgreen")

print(cummulative_plot)



#ggsave(cummulative_plot, filename = "cum_combined2.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")
```

The figure above is a typical of what I wanted. It shows the months where control should be focused to reduce the population of mosquitoes for different regions. From the Fig. above, we can see that the regions of Algeria and SA have more than half of the mosquito population for the entire year concentrated in two months. This implies that focusing controls in this months can significantly reduce the population and reduce disease spread.


## Follow up figure 
This figure explains the above figure by showing the month distribution for each region

```{r}
### --- This is just a trial code to attack the problem ---- ##

# data frame
df <- data.frame(region = rep("SA", 12),
                 month = month.abb,
                 rank = as.character(c(2, 4, 5, 8, 7, 10, 12, 1, 5, 3, 6, 9)),
                 region2 = rep("Gh", 12),
                 rank2 = as.character(c(5, 6, 8, 9, 2, 7, 1, 3, 4, 10, 12, 11)))

# Convert the "rank" column to numeric
df$rank <- as.numeric(df$rank)
df$rank2 <- as.numeric(df$rank2)

# Reorder the levels of the "month" factor
df$month <- factor(df$month, levels = month.abb)

# Create a combined point plot with rotated axes
ggplot(df, aes(x = month)) +
  geom_point(aes(y = factor(region), color = "SA"), shape = 0, size = 5) +  # Use shape 16 (filled square) for SA
  geom_point(aes(y = factor(region2), color = "Gh"), shape = 0, size = 5) +   # Use shape 1 (hollow square) for Gh
  labs(title = "Empty Square Point Plot of Ranks for SA and Gh Regions",
       x = "Month",
       y = "Region") +
  scale_color_manual(values = c("SA" = "black", "Gh" = "black")) +  # Set color for SA and Gh
  theme_bw() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1)) +
  guides(color = FALSE)  # Remove the color legend


# Create a combined point plot with rotated axes and a color legend
ggplot(df, aes(x = month)) +
  geom_point(aes(y = factor(region), color = rank), shape = 15, size = 5) +  # Use shape 15 (square) for SA
  geom_point(aes(y = factor(region2), color = rank2), shape = 15, size = 5) +  # Use shape 15 (square) for Gh
  geom_text(aes(y = factor(region), label = rank), vjust = 1.5, size = 3.5) +  # Add text labels for SA
  geom_text(aes(y = factor(region2), label = rank2), vjust = 1.5, size = 3.5) +  # Add text labels for Gh
  labs(title = "Heatmap of Ranks for SA and Gh Regions",
       x = "Month",
       y = "Region") +
  scale_color_gradient(low = "lightblue", high = "darkblue", name = "Rank") +  # Set color scale
  theme_bw() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1)) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 1, size = 11),  # Adjust size as needed
    axis.text.y = element_text(size = 11),  # Adjust size as needed
    axis.title = element_text(size = 12),
    legend.text = element_text(size = 11)# Adjust size as needed
  )
```


```{r}
df.alg <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_alg.csv")
df.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")
df.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")
df.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")

# Create a column called count to data which is same as total adult numbers
df.alg <- df.alg %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
algeria_data <- df.alg %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Create a vector of months of the year
months_list <- month.abb

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
algeria_data <- algeria_data %>%
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
algeria_sorted_data <- algeria_data[order(-algeria_data$monthly_percentage),]

# Save the reordered data as a new data frame
algeria_data <- data.frame(months_alg = factor(algeria_sorted_data$months),
                       ranked_months = factor(1:12),
                       monthly_percentage_alg = algeria_sorted_data$monthly_percentage)

# Define the month order using the constant month.abb
month_order <- month.abb

# Reorder the data frame based on the month order
alg_ordered_data <- algeria_data[match(month_order, algeria_data$months_alg), ]

alg_ordered_data <- alg_ordered_data %>% 
  rename(month = months_alg, rank_alg = ranked_months) %>% 
  mutate(region_alg = rep("ALGR"))

## ---------------------------------------- ----------------------------------------
## Ghana follows here
## ---------------------------------------- ----------------------------------------
 
# Create a column called count to data which is same as total adult numbers
df.gh <- df.gh %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
ghana_data <- df.gh %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
ghana_data <- ghana_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
ghana_sorted_data <- ghana_data[order(-ghana_data$monthly_percentage),]

# Save the reordered data as a new data frame
ghana_data <- data.frame(months_gh = factor(ghana_sorted_data$months),
                       ranked_months = factor(1:12),
                       monthly_percentage_gh = ghana_sorted_data$monthly_percentage)

# Reorder the data frame based on the month order
gh_ordered_data <- ghana_data[match(month_order, ghana_data$months_gh), ]

gh_ordered_data <- gh_ordered_data %>% 
  rename(month = months_gh, rank_gh = ranked_months) %>% 
  mutate(region_gh = rep("GHAN"))

## ---------------------------------------- ----------------------------------------
## Kenya follows here
## ---------------------------------------- ----------------------------------------
# Create a column called count to data which is same as total adult numbers
df.kenya <- df.kenya %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
kenya_data <- df.kenya %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
kenya_data <- kenya_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
kenya_sorted_data <- kenya_data[order(-kenya_data$monthly_percentage),]

# Save the reordered data as a new data frame
kenya_data <- data.frame(months_kenya = factor(kenya_sorted_data$months),
                       ranked_months = factor(1:12),
                       monthly_percentage_kenya = kenya_sorted_data$monthly_percentage)

# Reorder the data frame based on the month order
kenya_ordered_data <- kenya_data[match(month_order, kenya_data$months_kenya), ]

kenya_ordered_data <- kenya_ordered_data %>% 
  rename(month = months_kenya, rank_kenya = ranked_months) %>% 
  mutate(region_kenya = rep("KENY"))

## ---------------------------------------- ----------------------------------------
## SA section here
## ---------------------------------------- ----------------------------------------
# Create a column called count to data which is same as total adult numbers
df.sa <- df.sa %>% 
  mutate(count = vary.total.adults)

# Extract the month information from the time column, then calculate the monthly percentages
sa_data <- df.sa %>%
  mutate(month = as.integer(format(as.Date(time, origin="2021-01-01"), "%m"))) %>%
  group_by(month) %>%
  summarise(monthly_total = sum(count)) %>%
  mutate(monthly_percentage = (monthly_total / sum(monthly_total)) * 100)

# Add a column of months to the data and arrange months in decreasing order of monthly percentages
sa_data <- sa_data %>% 
  mutate(months = factor(months_list, levels = months_list[order(monthly_percentage, decreasing = TRUE)]))

# Reorder the data frame based on decreasing fraction
sa_sorted_data <- sa_data[order(-sa_data$monthly_percentage),]

# Save the reordered data as a new data frame
sa_data <- data.frame(months_sa = factor(sa_sorted_data$months),
                       ranked_months = factor(1:12),
                       monthly_percentage_sa = sa_sorted_data$monthly_percentage)

# Reorder the data frame based on the month order
sa_ordered_data <- sa_data[match(month_order, sa_data$months_sa), ]

sa_ordered_data <- sa_ordered_data %>% 
  rename(month = months_sa, rank_sa = ranked_months) %>% 
  mutate(region_sa = rep("SA"))

## ---------------------------------------- ----------------------------------------
## Merge data and proceed to plot
## ---------------------------------------- ----------------------------------------
rank_data <- inner_join(alg_ordered_data, gh_ordered_data) %>% 
  inner_join(kenya_ordered_data) %>% 
  inner_join(sa_ordered_data)

# Convert the "rank" column to numeric
rank_data$rank_alg <- as.numeric(rank_data$rank_alg)
rank_data$rank_gh <- as.numeric(rank_data$rank_gh)
rank_data$rank_kenya <- as.numeric(rank_data$rank_kenya)
rank_data$rank_sa <- as.numeric(rank_data$rank_sa)

# Reorder the levels of the "month" factor
rank_data$month <- factor(rank_data$month, levels = month.abb)

# Create a point plot with rotated axes and a color legend
ggplot(rank_data, aes(x = month)) +
  geom_point(aes(y = factor(region_alg), color = rank_alg), shape = 15, size = 8) +  # Use shape 15 (square) for Algeria
  geom_point(aes(y = factor(region_gh), color = rank_gh), shape = 15, size = 8) +  # Use shape 15 (square) for Ghana
  geom_point(aes(y = factor(region_kenya), color = rank_kenya), shape = 15, size = 8) +  # Use shape 15 (square) for Ghana
  geom_point(aes(y = factor(region_sa), color = rank_sa), shape = 15, size = 8) +  # Use shape 15 (square) for Ghana
  geom_text(aes(y = factor(region_alg), label = rank_alg), vjust = 1, size = 4) +  # Add text labels for SA
  geom_text(aes(y = factor(region_gh), label = rank_gh), vjust = 1, size = 4) +  # Add text labels for Gh
  geom_text(aes(y = factor(region_kenya), label = rank_kenya), vjust = 1, size = 4) +  # Add text labels for SA
  geom_text(aes(y = factor(region_sa), label = rank_sa), vjust = 1, size = 4) +  # Add text labels for Gh
  labs(title = "Heatmap of Ranks for All the Regions",
       x = "Month",
       y = "Region") +
  scale_color_gradient(low = "darkblue", high = "lightblue", name = "Rank") +  # Set color scale
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

# Same plot as above with an arrow added
rank <- ggplot(rank_data, aes(x = month)) +
  geom_point(aes(y = factor(region_alg), color = rank_alg), shape = 15, size = 8, alpha = 0.7) +
  geom_point(aes(y = factor(region_gh), color = rank_gh), shape = 15, size = 8, alpha = 0.7) +
  geom_point(aes(y = factor(region_kenya), color = rank_kenya), shape = 15, size = 8, alpha = 0.7) +
  geom_point(aes(y = factor(region_sa), color = rank_sa), shape = 15, size = 8, alpha = 0.7) +
  geom_text(aes(y = factor(region_alg), label = rank_alg), vjust = 0.5, hjust = 0.5, size = 6) +
  geom_text(aes(y = factor(region_gh), label = rank_gh), vjust = 0.5, hjust = 0.5, size = 6) +
  geom_text(aes(y = factor(region_kenya), label = rank_kenya), vjust = 0.5, hjust = 0.5, size = 6) +
  geom_text(aes(y = factor(region_sa), label = rank_sa), vjust = 0.5, hjust = 0.5, size = 6) +
  labs(title = "Heatmap of ranks for the regions",
       x = "",
       y = "Region") +
  scale_color_gradient(low = "darkblue", high = "lightblue", name = "Rank") +
  scale_alpha_continuous(range = c(0.3, 1)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 12),
  )

rank

#ggsave(rank, filename = "rank.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")




```


```{r}
library(cowplot)
library(magick)

# Load saved figures
plot_cummulative <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/cum_combined2.png", scale = 1)
plot_rank <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/rank.png", scale = 1)

# Numbering the plots

numbered_plot_cummulative <- plot_cummulative + draw_label("(a)", x = 0, y = 1, vjust = 6, hjust = 0, size = 14) # Adjust hjust to move closer to the figure
numbered_plot_rank <- plot_rank + draw_label("(b)", x = 0, y = 1, vjust = 6, hjust = 0, size = 14) # Adjust hjust to move closer to the figure

# Combine the plots with reduced margins
combined_plot <- plot_grid(
  numbered_plot_cummulative, 
  numbered_plot_rank, 
  ncol = 2,
  rel_widths = c(1, 1),  # Adjust the relative widths of columns
  rel_heights = c(1)  # Adjust the relative heights of rows
) +
theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))  # Adjust the margins

# Display the combined plot
print(combined_plot)

#ggsave(combined_plot, filename = "cum_combined.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```


```{r}
cummulative_plot <- ggplot(merged_data, aes(x = ranked_months, group = 1)) +
  geom_step(aes(y = cum_fraction_alg, color = "Algeria"), size = 1.2, color = "purple") +
  geom_step(aes(y = cum_fraction_gh, color = "Ghana"), size = 1.2, color = "darkred") +
  geom_step(aes(y = cum_fraction_kenya, color = "Kenya"), size = 1.2, color = "orange") +
  geom_step(aes(y = cum_fraction_sa, color = "South Africa"), size = 1.2, color = "darkgreen") +
  ylab("Cumulative fraction of mosquitoes (%)") +
  xlab("Months up to rank") +
  ggtitle("") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 1, size = 11),  # Adjust size as needed
    axis.text.y = element_text(size = 11),  # Adjust size as needed
    axis.title = element_text(size = 12),
    legend.position = "none"  # Remove legend
  ) +
  ylim(0, max(merged_data$cum_fraction_alg)) +  # Set y-axis limits from 0 to the maximum cumulative fraction
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", size = 1.2) +
  geom_text(data = merged_data, 
            aes(x = max(as.numeric(levels(ranked_months))), 
                y = tail(cum_fraction_alg, 1),
                label = "ALGR"),
            vjust = -0.5, hjust = 6, color = "purple") +
  geom_text(data = merged_data, 
            aes(x = max(as.numeric(levels(ranked_months))), 
                y = tail(cum_fraction_gh, 1),
                label = "GHAN"),
            vjust = 4, hjust = 5, color = "darkred") +
  geom_text(data = merged_data, 
            aes(x = max(as.numeric(levels(ranked_months))), 
                y = tail(cum_fraction_kenya, 1),
                label = "KENY"),
            vjust = 6, hjust = 2, color = "orange") +
  geom_text(data = merged_data, 
            aes(x = max(as.numeric(levels(ranked_months))), 
                y = tail(cum_fraction_sa, 1),
                label = "SA"),
            vjust = 5.4, hjust = 20, color = "darkgreen")


# Same plot as above with an arrow added
rank <- ggplot(rank_data, aes(x = month)) +
  geom_point(aes(y = factor(region_alg), color = rank_alg), shape = 15, size = 8, alpha = 0.7) +
  geom_point(aes(y = factor(region_gh), color = rank_gh), shape = 15, size = 8, alpha = 0.7) +
  geom_point(aes(y = factor(region_kenya), color = rank_kenya), shape = 15, size = 8, alpha = 0.7) +
  geom_point(aes(y = factor(region_sa), color = rank_sa), shape = 15, size = 8, alpha = 0.7) +
  geom_text(aes(y = factor(region_alg), label = rank_alg), vjust = 0.5, hjust = 0.5, size = 6) +
  geom_text(aes(y = factor(region_gh), label = rank_gh), vjust = 0.5, hjust = 0.5, size = 6) +
  geom_text(aes(y = factor(region_kenya), label = rank_kenya), vjust = 0.5, hjust = 0.5, size = 6) +
  geom_text(aes(y = factor(region_sa), label = rank_sa), vjust = 0.5, hjust = 0.5, size = 6) +
  labs(title = "Heatmap of ranks for the regions",
       x = "",
       y = "Region") +
  scale_color_gradient(low = "darkblue", high = "lightblue", name = "Rank") +
  scale_alpha_continuous(range = c(0.3, 1)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 11),
    axis.text.y = element_text(size = 11),
    axis.title = element_text(size = 12),
  )

# Arranging the plots in a grid using facet_wrap
combined_plots <- cowplot::plot_grid(cummulative_plot, rank, ncol = 2)

# Print the combined plot
print(combined_plots)
```




## Plot of the fraction of larvae to adult mosquitoes for each month across the year.

Here, I try out code to attempt to produce the required plot. 

```{r}

#####-----------------------------------------------------------------------------
# Plot fraction of adults
#####-----------------------------------------------------------------------------
library(rio)

df.alg <- import("./data/df_algeria.csv")
df.gh <- import("./data/df_ghana.csv")
df.kenya <- import("./data/df_kenya.csv")
df.sa <- import("./data/df_SA.csv")

# Calculate the fraction of adult population at each time point. Algeria data
total_population <- rowSums(df.alg[,c("eggs", "larvae", "pupae", "total.adults")])
#fraction_adult <- df.alg[, "total.adults"] / total_population

df.alg <- df.alg %>% 
  mutate(fraction_adult = total.adults/total_population)

# Ghana data
total_pop.gh <- rowSums(df.gh[,c("eggs", "larvae", "pupae", "total.adults")])

df.gh <- df.gh %>% 
  mutate(fraction_adult = total.adults/total_pop.gh)

# Kenya data
total_pop.kenya <- rowSums(df.kenya[,c("eggs", "larvae", "pupae", "total.adults")])

df.kenya <- df.kenya %>% 
  mutate(fraction_adult = total.adults/total_pop.kenya)

# SA data
# Kenya data
total_pop.SA <- rowSums(df.sa[,c("eggs", "larvae", "pupae", "total.adults")])

df.sa <- df.sa %>% 
  mutate(fraction_adult = total.adults/total_pop.SA)


#-------------------------------------------------------------------------------
# Plot fraction of population that are adults
# -----------------------------------------------------------------------------
frac.adult <- data.frame(time = df.alg$time, Algeria=df.alg$fraction_adult, Ghana=df.gh$fraction_adult, Kenya=df.kenya$fraction_adult, South.Africa=df.sa$fraction_adult)


frac.adult <- frac.adult %>% 
  pivot_longer(cols=c("Algeria", "Ghana", "Kenya", "South.Africa"),
               names_to='Region',
               values_to='population')

ggp <- ggplot(frac.adult, aes(x=time, y=population, col = Region))+
  geom_line(size = 1)+
  xlab("Time (days)") +
  ylab("Fraction of population") +
  ggtitle("Fraction of female adults in different regions") +
  theme_classic()

ggp +                               # Set origin of axes to zero
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))

# At this point, I will change time on x-axis from days to months

frac.adult <- frac.adult %>% 
  rename(
    x = time,
    y = population
  )

breakfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  days <- origin + x
  origin <- as.POSIXlt(origin)
  dayseq <- as.POSIXlt(seq(days[1], days[2], by = "day"))
  with(dayseq, yday[mday == 1] + 365*(year[mday == 1] - origin$year ))
}

labelfunc <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

ggp2 <- ggplot(frac.adult, aes(x, y, col = Region)) +
  geom_line(size = 1) +
  labs(title = "Fraction of adult mosquitoes in different regions",
       x = "",
       y = "Fraction of adult mosquitoes") +
  theme_classic() +
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 60,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 12),
        text = element_text(size = 16)) +
  scale_x_continuous(breaks=breakfunc, labels=labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))

ggp2
# Save the plot
#ggsave("fraction_abundance.png", ggp2, path = "/Users/jbaafi/Desktop")


###########=----------------------------------------
# Comparing stage fractions
# ----------------------------------------
# Extract the number of mosquitoes in each stage
E <- df.alg[, "eggs"]
L <- df.alg[, "larvae"]
P <- df.alg[, "pupae"]
A <- df.alg[, "total.adults"]

# Calculate the total number of mosquitoes at each time step
N <- E + L + P + A

# Calculate the fraction of mosquitoes in each stage
frac_E <- E/N
frac_L <- L/N
frac_P <- P/N
frac_A <- A/N

df.alg <- df.alg %>% 
  mutate(frac_E, frac_L, frac_P, frac_A)


colors <- c("Larvae" = "blue", "Adult" = "red")

ggp.alg <- ggplot(df.alg, aes(x = time)) +
  geom_line(aes(y = frac_L, color = "Larvae"), size = 1) +
  geom_line(aes(y = frac_A, color = "Adult"), size = 1) +
  labs(x = "",
       y = "Fraction of population",
       color = "Legend", 
       title = "Fraction of population in Algeria") +
  scale_color_manual(values = colors) +
  theme_classic()+
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 60,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 12),
        text = element_text(size = 16)) +
  scale_x_continuous(breaks=breakfunc, labels=labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))

ggp.alg

#. ---------------------------------
# Fraction for Ghana
# ----------------------
# Extract the number of mosquitoes in each stage
E <- df.gh[, "eggs"]
L <- df.gh[, "larvae"]
P <- df.gh[, "pupae"]
A <- df.gh[, "total.adults"]

# Calculate the total number of mosquitoes at each time step
N <- E + L + P + A

# Calculate the fraction of mosquitoes in each stage
frac_E <- E/N
frac_L <- L/N
frac_P <- P/N
frac_A <- A/N

df.gh <- df.gh %>% 
  mutate(frac_E, frac_L, frac_P, frac_A)


colors <- c("Larvae" = "blue", "Adult" = "red")

ggp.gh <- ggplot(df.gh, aes(x = time)) +
  geom_line(aes(y = frac_L, color = "Larvae"), size = 1) +
  geom_line(aes(y = frac_A, color = "Adult"), size = 1) +
  labs(x = "",
       y = "Fraction of population",
       color = "Legend", 
       title = "Fraction of population in Ghana") +
  scale_color_manual(values = colors) +
  theme_classic()+
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 60,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 12),
        text = element_text(size = 16)) +
  scale_x_continuous(breaks=breakfunc, labels=labelfunc, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))

ggp.gh

```


The chunk below show the fraction of larvae that emerge as adults. 

This plot is great to demonstrate how temperature and rainfall influence the fraction of larvae that emerge as adults over time. The rate of emergence is highly dependent on the combined effects of temperature and rainfall. 
Annotations are added to see the events that influence the fraction of larvae that emerge as adults. The pattern seen is consistent with theory. 

```{r}
library(rio)
# This does prety same as before
df.alg <- import("./data/df_algeria.csv")

df.alg <- df.alg[1461:nrow(df.alg), ]

mosquito_data <- df.alg
mosquito_data$frac_emerge_adults <- mosquito_data$total.adults / mosquito_data$larvae

# Modify the time column to start from 1
mosquito_data$time <- seq_along(mosquito_data$time)

# Temperature Plot
p1 <- ggplot(mosquito_data, aes(x = time, y = temperature)) +
  geom_line() +
  labs(title = "Algeria",
       x = "",
       y = bquote("Temperature (" ~ degree ~ "C)")) +
  theme_classic()

# Rainfall Plot
p2 <- ggplot(mosquito_data, aes(x = time, y = rainfall)) +
  geom_line() +
  labs(title = "",
       x = "",
       y = "Rainfall (mm)") +
  theme_classic()


# Fraction of Larvae Emerging as Adults Plot
p3 <- ggplot(mosquito_data, aes(x = time, y = frac_emerge_adults)) +
  geom_line(col = "purple", lwd = 1.5) +
  labs(title = "",
       x = "Time (days)",
       y = "Fraction of larvae emerging as adults") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 12),
    axis.text.y = element_text(colour = "grey20", size = 12),
    text = element_text(size = 13)) +
  ggtitle("Algeria")+
  theme(plot.title = element_text(hjust = 0.5, size = 16))

# Annotate low and high temperature events
low_temp_annotation <- annotate("text", x = 22, y = 0.15, label = "Low temp ", color = "red", size = 5)
high_temp_annotation <- annotate("text", x = 204, y = 0.85, label = "High temp", color = "blue", size = 5)

# Annotate low and high rainfall events
low_rainfall_annotation <- annotate("text", x = 139, y = 0.1, label = "Low rainfall", color = "red", size = 5)
high_rainfall_annotation <- annotate("text", x = 217, y = 0.8, label = "High rainfall", color = "blue", size = 5)

# Combine the base plot and annotations
p3 <- p3 + low_temp_annotation + high_temp_annotation + low_rainfall_annotation + high_rainfall_annotation

library(gridExtra)

# Combine Plots
combined_plots <- grid.arrange(
  p1, p2, p3,  # Replace p1, p2, p3 with your actual plot objects
  ncol = 1  # Number of columns in the combined plot
)

p3
ggsave(p3, filename = "larv_emerge_Algeria2.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")


```
The plot below focuses on the proportion of mosquito that are larvae for each region. 
It is inspired by Amy's sea lice paper. Refer to that paper for more indept explanation of this fig. 




```{r}
library(rio)
# This does pretty same as before but utilizes data from Ghana
df.gh <- import("./data/df_ghana.csv")

df.gh <- df.gh[1461:nrow(df.gh), ]

mosquito_data_gh <- df.gh
mosquito_data_gh$frac_emerge_adults <- mosquito_data_gh$total.adults / mosquito_data_gh$larvae

# Modify the time column to start from 1
mosquito_data_gh$time <- seq_along(mosquito_data_gh$time)

# Temperature Plot
p1.gh <- ggplot(mosquito_data_gh, aes(x = time, y = temperature)) +
  geom_line() +
  labs(title = "Ghana",
       x = "",
       y = bquote("Temperature (" ~ degree ~ "C)"))+
  theme_classic()

# Rainfall Plot
p2.gh <- ggplot(mosquito_data_gh, aes(x = time, y = rainfall)) +
  geom_line() +
  labs(title = "",
       x = "",
       y = "Rainfall (mm)") +
  theme_classic()

# Fraction of Larvae Emerging as Adults Plot
p3.gh <- ggplot(mosquito_data_gh, aes(x = time, y = frac_emerge_adults)) +
  geom_line(col = "darkred", lwd = 1.5) +
  labs(title = " ",
       x = "Time (days)",
       y = "Fraction of larvae emerging as adults") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 12),
    axis.text.y = element_text(colour = "grey20", size = 12),
    text = element_text(size = 13)) +
  ggtitle("Ghana")+
  theme(plot.title = element_text(hjust = 0.5, size = 16))

# Annotate low and high temperature events
low_temp_annotation <- annotate("text", x = 235, y = 0.33, label = "Low temp", color = "red", size = 5)
high_temp_annotation <- annotate("text", x = 52, y = 0.48, label = "High temp", color = "blue", size = 5)

# Annotate low and high rainfall events
low_rainfall_annotation <- annotate("text", x = 17, y = 0.42, label = "Low rainfall", color = "red", size = 5)
high_rainfall_annotation <- annotate("text", x = 200, y = 0.35, label = "High rainfall", color = "blue", size = 5)

# Combine the base plot and annotations
p3.gh <- p3.gh + low_temp_annotation + high_temp_annotation + low_rainfall_annotation + high_rainfall_annotation

p3.gh
ggsave(p3.gh, filename = "larv_emerge_ghana2.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")
```



```{r}
library(rio)
library(ggplot2)
# This does pretty same as before but utilizes data from SA
df.sa <- import("./data/df_SA.csv")

df.sa <- df.sa[1461:nrow(df.sa), ]

mosquito_data_sa <- df.sa
mosquito_data_sa$frac_emerge_adults <- mosquito_data_sa$total.adults / mosquito_data_sa$larvae

# Modify the time column to start from 1
mosquito_data_sa$time <- seq_along(mosquito_data_sa$time)

# Temperature Plot
p1.sa <- ggplot(mosquito_data_sa, aes(x = time, y = temperature)) +
  geom_line() +
  labs(title = "South Africa",
       x = "",
       y = bquote("Temperature (" ~ degree ~ "C)")) +
  theme_classic()

# Rainfall Plot
p2.sa <- ggplot(mosquito_data_sa, aes(x = time, y = rainfall)) +
  geom_line() +
  labs(y = "Rainfall (mm)",
       x = "") +
  theme_classic()

# Fraction of Larvae Emerging as Adults Plot
p3.sa <- ggplot(mosquito_data_sa, aes(x = time, y = frac_emerge_adults)) +
  geom_line(col = "darkgreen", lwd = 1.5) +
  labs(title = " ",
       x = "Time (days)",
       y = "Fraction of larvae emerging as adults") +
  theme_classic() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 12),
    axis.text.y = element_text(colour = "grey20", size = 12),
    text = element_text(size = 13))+
  ggtitle("South Africa")+
  theme(plot.title = element_text(hjust = 0.5, size = 16))

# Find the row index where temperature is minimum
min_row_temp <- which.min(mosquito_data_sa$temperature)

# Access the time for the minimum rainfall
time_for_min_temp <- mosquito_data_sa$time[min_row_temp]

# Maximum temperature
max_row_temp <- which.max(mosquito_data_sa$temperature)
time_for_max_temp <- mosquito_data_sa$time[max_row_temp]

# Annotate low and high temperature events
low_temp_annotation <- annotate("text", x = time_for_min_temp, y = 0.1, label = "Low temp", color = "red", size = 5)
high_temp_annotation <- annotate("text", x = time_for_max_temp, y = 0.15, label = "High temp", color = "blue", size = 5)

# Find the row index where rainfall is minimum
min_row_rain <- which.min(mosquito_data_sa$rainfall)

# Access the time for the minimum rainfall
time_for_min_rain <- mosquito_data_sa$time[min_row_rain]

# Maximum rainfall
max_row_rain <- which.max(mosquito_data_sa$rainfall)
time_for_max_rain <- mosquito_data_sa$time[max_row_rain]

# Annotate low and high rainfall events
low_rainfall_annotation <- annotate("text", x = time_for_min_rain, y = 0.1, label = "Low rainfall", color = "red", size = 4.2)
high_rainfall_annotation <- annotate("text", x = time_for_max_rain, y = 0.15, label = "High rainfall", color = "blue", size = 5)

# Combine the base plot and annotations
p3.sa <- p3.sa + low_temp_annotation + high_temp_annotation + low_rainfall_annotation + high_rainfall_annotation

library(gridExtra)

# Combine Plots
combined_plots_sa <- grid.arrange(
  p1.sa, p2.sa, p3.sa,  # Replace p1, p2, p3 with your actual plot objects
  ncol = 1  # Number of columns in the combined plot
)

p3.sa

ggsave(p3.sa, filename = "larv_emerge_sa2.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")
```


 
```{r}
library(rio)
library(ggplot2)
# This does pretty same as before but utilizes data from SA
df.kenya <- import("./data/df_kenya.csv")

df.kenya <- df.kenya[1461:nrow(df.kenya), ]

mosquito_data_kenya <- df.kenya
mosquito_data_kenya$frac_emerge_adults <- mosquito_data_kenya$total.adults / mosquito_data_kenya$larvae

# Modify the time column to start from 1
mosquito_data_kenya$time <- seq_along(mosquito_data_kenya$time)

# Temperature Plot
p1.kenya <- ggplot(mosquito_data_kenya, aes(x = time, y = temperature)) +
  geom_line() +
  labs(title = "Kenya",
       x = "",
       y = bquote("Temperature (" ~ degree ~ "C)")) +
  theme_classic()

# Rainfall Plot
p2.kenya <- ggplot(mosquito_data_kenya, aes(x = time, y = rainfall)) +
  geom_line() +
  labs(y = "Rainfall (mm)",
       x = "") +
  theme_classic()

# Fraction of Larvae Emerging as Adults Plot
p3.kenya <- ggplot(mosquito_data_kenya, aes(x = time, y = frac_emerge_adults)) +
  geom_line(col = "orange", lwd = 1.5) +
  labs(title = " ",
       x = "Time (days)",
       y = "Fraction of larvae emerging as adults") +
  theme_classic()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 12),
    axis.text.y = element_text(colour = "grey20", size = 12),
    text = element_text(size = 13))+
  ggtitle("Kenya")+
  theme(plot.title = element_text(hjust = 0.5, size = 16))

# Find the row index where temperature is minimum
min_row_temp <- which.min(mosquito_data_kenya$temperature)

# Access the time for the minimum rainfall
time_for_min_temp <- mosquito_data_kenya$time[min_row_temp]

# Maximum temperature
max_row_temp <- which.max(mosquito_data_kenya$temperature)
time_for_max_temp <- mosquito_data_kenya$time[max_row_temp]

# Annotate low and high temperature events
low_temp_annotation <- annotate("text", x = time_for_min_temp, y = 0.3, label = "Low temp", color = "red", size = 5)
high_temp_annotation <- annotate("text", x = time_for_max_temp, y = 0.53, label = "High temp", color = "blue", size = 5)

# Find the row index where rainfall is minimum
min_row_rain <- which.min(mosquito_data_kenya$rainfall)

# Access the time for the minimum rainfall
time_for_min_rain <- mosquito_data_kenya$time[min_row_rain]

# Maximum rainfall
max_row_rain <- which.max(mosquito_data_kenya$rainfall)
time_for_max_rain <- mosquito_data_kenya$time[max_row_rain]

# Annotate low and high rainfall events
low_rainfall_annotation <- annotate("text", x = time_for_min_rain, y = 0.44, label = "Low rainfall", color = "red", size = 5)
high_rainfall_annotation <- annotate("text", x = time_for_max_rain, y = 0.32, label = "High rainfall", color = "blue", size = 5)

# Combine the base plot and annotations
p3.kenya <- p3.kenya + low_temp_annotation + high_temp_annotation + low_rainfall_annotation + high_rainfall_annotation

library(gridExtra)

# Combine Plots
combined_plots_kenya <- grid.arrange(
  p1.kenya, p2.kenya, p3.kenya,  # Replace p1, p2, p3 with your actual plot objects
  ncol = 1  # Number of columns in the combined plot
)

p3.kenya

ggsave(p3.kenya, filename = "larv_emerge_kenya2.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")
```


The plot from SA and Algeria clearly demonstrate the effect of temperature and rainfall on mosquito larvae that develop to emerge as adults. 
The plot for Ghana does not give a clear demonstration of this concept since temperature stays high across the year but it is fairly good.


Still the main plot I wanted to do is not done yet. Continue to do the plot for the fraction of population that are adults and compare to the larvae pop.


## A few notes:

I think that the plot that investigates the fraction of larvae that emerge as adults over time gives a great insight into optimizing control strategies. The basic interpretation will be that adult control may have to be intensified in the periods where higher fraction of the larvae emerge as adult. Also much attention may be given to the period where only a small fraction of the larvae emerge as adult. In this case larvicides may be the best optimal control strategy to be implemented. 


## Another section that plots total number of adult across the year

This chunk only selects and plots the 5th year total and plots it.
```{r}
# Load data from various regions 
df.alg <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_alg.csv")
df.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")
df.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")
df.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")

data <- data.frame(time = df.alg$time, 
                   algeria = df.alg$vary.total.adults, 
                   ghana = df.gh$vary.total.adults, 
                   kenya = df.kenya$vary.total.adults, 
                   sa = df.sa$vary.total.adults)


data <- data[1461:nrow(data), ]

# Calculate total adults for each region at the final time point
total_adults <- data.frame(
  Region = c("Algeria", "South Africa", "Kenya", "Ghana"),
  total = c(sum(data$algeria), sum(data$sa), sum(data$kenya), sum(data$ghana))
)

# Create the bar plot without legend
total <- ggplot(total_adults, aes(x = Region, y = total, fill = Region)) +
  geom_bar(stat = "identity", color = "black", width = 0.3, alpha = 0.85) +
  scale_fill_manual(values = c("Algeria" = "yellow", "South Africa" = "blue", "Kenya" = "orange", "Ghana" = "red")) +
  theme_classic() + 
  scale_y_continuous(bquote("Mosquito abundance" ~ (x ~ 10^5)), breaks = seq(0, 5, 1) * 1e5, labels = seq(0, 5, 1)) +
  theme(legend.position = "none") +
  ggtitle("Estimated total number of mosquitoes")

total

#ggsave(total, filename = "total_adults.png", path = "/Users/jbaafi/Desktop/plots")

```


This chunk combines some saved figures into rows and columns to be included in the manuscript
```{r}
library(cowplot)
library(magick)

# Load your saved figures
# Load your saved figures

plot_adult <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/adult_abundance.png", scale = 1)
plot_total <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/total_adults.png", scale = 1)
plot_algeria <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/larv_emerge_Algeria2.png", scale = 1)
plot_sa <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/larv_emerge_sa2.png", scale = 1)
plot_kenya <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/larv_emerge_kenya2.png", scale = 1)
plot_ghana <- ggdraw() + draw_image("/Users/jbaafi/Documents/Github/research-project-one/saved plots/larv_emerge_ghana2.png", scale = 1)

# Numbering the plots

numbered_plot_adult <- plot_adult + draw_label("(a)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14) 
numbered_plot_total <- plot_total + draw_label("(b)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14)
numbered_plot_algeria <- plot_algeria + draw_label("(c)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14)
numbered_plot_sa <- plot_sa + draw_label("(d)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14)
numbered_plot_kenya <- plot_kenya + draw_label("(e)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14)
numbered_plot_ghana <- plot_ghana + draw_label("(f)", x = 0, y = 1, vjust = 1, hjust = 0, size = 14)



# Combine the plots
combined_plot <- plot_grid(
  numbered_plot_adult, 
  numbered_plot_total, 
  numbered_plot_algeria, 
  numbered_plot_sa, 
  numbered_plot_kenya, 
  numbered_plot_ghana, 
  ncol = 2
)

# Display the combined plot
print(combined_plot)

ggsave(combined_plot, filename = "Fig3_combined.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```



This chunk plots the average of the 5 years and plot it. Similar to the previous except that here, we are plotting 5 year average.
```{r}
# Load data from various regions 
df.alg <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_alg.csv")
df.gh <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_gh.csv")
df.kenya <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_kenya.csv")
df.sa <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/numerical_data/variable.condt_sa.csv")

df.alg <- df.alg %>% 
  mutate(count = vary.total.adults)

df.gh <- df.gh %>% 
  mutate(count = vary.total.adults)

df.kenya <- df.kenya %>% 
  mutate(count = vary.total.adults)

df.sa <- df.sa %>% 
  mutate(count = vary.total.adults)

library(lubridate)

mosquito_data.alg <- df.alg %>%
  mutate(date = as.Date(time, origin = "2019-01-01"),
         year = year(date),
         month = month(date)) %>%
  group_by(year) %>%
  summarise(yearly_total = sum(count))

mosquito_data.gh <- df.gh %>%
  mutate(date = as.Date(time, origin = "2019-01-01"),
         year = year(date),
         month = month(date)) %>%
  group_by(year) %>%
  summarise(yearly_total = sum(count))

mosquito_data.kenya <- df.kenya %>%
  mutate(date = as.Date(time, origin = "2019-01-01"),
         year = year(date),
         month = month(date)) %>%
  group_by(year) %>%
  summarise(yearly_total = sum(count))

mosquito_data.sa <- df.sa %>%
  mutate(date = as.Date(time, origin = "2019-01-01"),
         year = year(date),
         month = month(date)) %>%
  group_by(year) %>%
  summarise(yearly_total = sum(count))


# Calculate total adults for each region at the final time point
df_average_mosquito <- data.frame(
  Region = c("Algeria", "South Africa", "Kenya", "Ghana"),
  average_count = c(mean(mosquito_data.alg$yearly_total), 
                    mean(mosquito_data.sa$yearly_total), 
                    mean(mosquito_data.kenya$yearly_total),
                    mean(mosquito_data.gh$yearly_total))
)

# Create the bar plot without legend
abundance <- ggplot(df_average_mosquito, aes(x = Region, y = average_count, fill = Region)) +
  geom_bar(stat = "identity", color = "black", width = 0.3, alpha = 0.85) +
  scale_fill_manual(values = c("Algeria" = "purple", "South Africa" = "darkgreen", "Kenya" = "orange", "Ghana" = "darkred")) +
  theme_classic() + 
  scale_y_continuous(bquote("Mosquito abundance" ~ (x ~ 10^5)), breaks = seq(0, 5, 1) * 1e5, labels = seq(0, 5, 1)) +
  theme(legend.position = "none") +
  ggtitle("Average Estimated Mosquito Abundance for Five Years")+
  theme(axis.text.x = element_text(colour = "grey20", size = 14, angle = 0,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 14),
        text = element_text(size = 16),
        plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete(labels = c("Algeria" = "ALGR",
                              "Ghana" = "GHAN",
                              "Kenya" = "KENY",
                              "South Africa" = "SA"))

abundance
#ggsave(abundance, filename = "total_adults.png", path = "/Users/jbaafi/Documents/Github/research-project-one/saved plots")

```






## Mosquito Count Data

```{r}
# The data comes from the Insect Control adult mosquito traps located in the City of Winnipeg and 10 km outside the city limits. The exact locations of the data have not been provided due to privacy issues. In the past, the file format has met FIPPA requirements. This data is updated daily.

daily_mosq_data <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/daily mosquito data/Daily_Adult_Mosquito_Trap_Data.csv")


mosq_larval_count_data <- read.csv("/Users/jbaafi/Documents/Github/research-project-one/daily mosquito data/Mosquito_Larval_Counts.csv")

ncol(daily_mosq_data)
colnames(daily_mosq_data)

```



